---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Flag ID

This chapter details the results of our flag analysis. 

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE}
#| label: set-chunk-options
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)

# Load packages
library(tidyverse)
library(kableExtra)

# Load data
flag_data <- read_csv("/capstone/seamissions/data/meds_meds_capstone_annual_vessel_flag_country_emissions_v20241121.csv", show_col_types = FALSE)
```

# About

The following analysis looks into the relationship between a vessel's AIS-registered flag and its top-visited (home port) country using data from emLab and Global Fishing Watch.

## Datasets

`meds_meds_capstone_annual_vessel_flag_country_emissions_v20241121.csv`

- `ssvid`: unique vessel ID
- `flag`: AIS-registered country
- `top_visited_country_iso3`: country the vessel visited the most number of times; home port
- `year`: year
- `emissions_co2_mt`: annual CO₂ emissions (MT)

## Packages

- {tidyverse}
- {kableExtra}

# Methods

## 1. Full Dataset Mismatch

To give a high level sense of just how much of an issue flags of convenience could be, we assessed AIS-registered flag and top-visited country mismatch for the dataset as a whole.

To do this, we created a `match` column, populated with the following values:

-   `TRUE`: match
-   `FALSE`: mismatch
-   `NA`: no value in AIS-flag, cannot determine.

```{r}
#| code-summary: Full Dataset Mismatch
# Investigate matching flags vs. non-matching flags
flag_match <- flag_data %>%
  # Fill `top_visit` NA values with `flag` values (**assumption**)
  mutate(top_visited_country_iso3 = ifelse(is.na(top_visited_country_iso3), flag, top_visited_country_iso3)) %>%
  mutate(match = case_when(flag == top_visited_country_iso3 ~ TRUE,
                           flag != top_visited_country_iso3 ~ FALSE))

# Summarise counts for TRUE match, FALSE mismatch, and NA
total_match <- flag_match %>%
  #filter(!is.na(flag)) %>%
  group_by(match) %>%
  summarise(counts = n()) %>%
  mutate(percent = round((counts/sum(counts))*100)) %>%
  select(-counts)

kbl(total_match, caption = "Flag Mismatch") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = F)
```

## 2. Full Dataset Emissions

Next, we looked at how much emissions (MT), on aggregate, could be affected by this flagging issue. We summarized emissions by match (TRUE), mismatch (FALSE), and NA for each year and overall.

```{r}
#| code-summary: Full Dataset Emissions
# Summarise emissions mismatch by year
year_emissions <- flag_match %>%
  #filter(!is.na(flag)) %>%
  group_by(year, match) %>%
  summarise(emissions = sum(emissions_co2_mt, na.rm = TRUE)) %>%
  group_by(year) %>%
  mutate(total_emissions = sum(emissions, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop = round(emissions/total_emissions*100, 1))

# Summarise emissions mismatch overall
year_emissions_avg <- year_emissions %>%
  group_by(match) %>%
  summarise(avg_prop = mean(prop),
            #check
            sum_emissions = sum(emissions)) %>%
  select(-avg_prop)

kbl(year_emissions_avg, caption = "Flag Mismatch Emissions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = F)
```

## 3. Overestimating Emissions

To quantify overestimation, or emissions attributed to a AIS-flag that visit different top-country, we assessed the fraction of emissions that end up in a different country (mismatch/FALSE) for each flag by year.

Ex. If for the flag of Panama (PAN) 75% of emissions are from vessels that have a different home port country, this could mean that we are *overestimating* emissions for Panama by upwards of 75% due to flagging issues.

```{r}
#| code-summary: Overestimating Emissions
# Summarise emissions attributed to mismatch by flag (AIS-flag)
overestimation_emissions <- flag_match %>%
  group_by(flag, match) %>%
  summarise(vessels = n(),
            emissions = sum(emissions_co2_mt, na.rm = TRUE)) %>%
  group_by(flag) %>%
  mutate(total_emissions = sum(emissions, na.rm = TRUE)) %>%
  mutate(percent_emissions_leaving_by_flag = round(emissions/total_emissions*100, 1)) %>%
  filter(match == FALSE) %>%
  ungroup()

overestimation_table <- overestimation_emissions %>%
  arrange(desc(emissions)) %>%
  slice(1:10) %>%
  select(-c(match, vessels, emissions))

kbl(overestimation_table, caption = "Overestimation of Emissions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = F)
```

## 4. Underestimating Emissions

To quantify underestimation, we assessed emissions for each top-visit country that come from a different AIS-flag.

Ex. If for the country of China 25% of emissions are from vessels that have a different AIS-flag, this could mean that we are *underestimating* emissions for China by upwards of 25% due to flagging issues.

```{r}
#| code-summary: Underestimating Emissions
# Summarise emissions attributed to mismatch by flag (top-country)
underestimation_emissions <- flag_match %>%
  group_by(top_visited_country_iso3, match) %>%
  summarise(vessels = n(),
            emissions = sum(emissions_co2_mt, na.rm = TRUE)) %>%
  group_by(top_visited_country_iso3) %>%
  mutate(total_emissions = sum(emissions, na.rm = TRUE)) %>%
  mutate(percent_emissions_coming_by_flag = round(emissions/total_emissions*100, 1)) %>%
  filter(match == FALSE) %>%
  rename(flag = top_visited_country_iso3) %>%
  ungroup()

# Look at CHN
# underestimation_emissions %>%
#   filter(flag == "CHN")

underestimation_table <- underestimation_emissions %>%
  arrange(desc(emissions)) %>%
  slice(1:10) %>%
  select(-c(match, vessels, emissions))

kbl(underestimation_table, caption = "Underestimation of Emissions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = F)
```

## Visualize

```{r}
#| code-summary: Visualize
# Top countries where emissions (prop) end up somewhere else (are flags of convenience)
# overestimation_emissions %>%
#   arrange(desc(percent_emissions_leaving_by_flag)) %>%
#   filter(percent_emissions_leaving_by_flag >= 99) %>%
# ggplot(aes(x = reorder(flag, percent_emissions_leaving_by_flag), y = percent_emissions_leaving_by_flag)) +
#   geom_col() +
#   coord_flip() +
#   labs(title = "Emissions Leaving (Overestimation)",
#     y = "Percent Emissions") +
#   theme_minimal() +
#   theme(
#     axis.title.y = element_blank()
#   )

# top countries where emissions (tons) end up somewhere else
overestimation_emissions %>%
  arrange(desc(emissions)) %>%
  slice(1:10) %>%
ggplot(aes(x = reorder(flag, emissions), y = emissions)) +
  geom_col() +
  coord_flip() +
  labs(title = "Emissions Leaving (Overestimation)",
       y = "Emissions (Metric Tons)") +
  theme_minimal() +
  theme(
    axis.title.y = element_blank()
  )

# top countries where emissions (prop) come from a different flag (utilize flags of convenience)
# underestimation_emissions %>%
#   arrange(desc(percent_emissions_coming_by_flag)) %>%
#   filter(percent_emissions_coming_by_flag >= 99) %>%
#   ggplot(aes(x = reorder(flag, percent_emissions_coming_by_flag), y = percent_emissions_coming_by_flag)) +
#   geom_col() +
#   coord_flip() +
#   labs(title = "Emissions Coming (Underestimation)",
#     y = "Percent Emissions") +
#   theme_minimal() +
#   theme(
#     axis.title.y = element_blank()
#   )

# top countries where emissions (tons) come from a different flag (utilize flags of convenience)
underestimation_emissions %>%
  arrange(desc(emissions)) %>%
  slice(1:10) %>%
ggplot(aes(x = reorder(flag, emissions), y = emissions)) +
  geom_col() +
  coord_flip() +
  labs(title = "Emissions Coming (Underestimation)",
       y = "Emissions (Metric Tons)") +
  theme_minimal() +
  theme(
    axis.title.y = element_blank()
  )
```

## Assumptions

Where top-visited country is NA, it was filled with AIS-flag value (assumes proper registration).

Vessels are assumed to land catch in their top-visited country, and that catch gets reported to the FAO by that country.