---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Merge Emissions Datasets {#sec-methods}

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE}
#| label: set-chunk-options
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)

library(tidyverse)

#tar_target(emissions, merge_emissions(emissions_files))
targets::tar_load(emissions)

source("R/functions.R")
```

## Datasets

Two emissions datasets, obtained from emLab, were used in this analysis:

- 1. Broadcasting emissions: `meds_capstone_ais_emissions_data_v20241121.csv`

- 2. Non-broadcasting emissions: `meds_capstone_non_broadcasting_emissions_data_v20250116.csv`

The data were pre-filtered by emLab to select for fishing vessels. The following columns are needed in the analysis:

- `month`
- `flag`
- `vessel_class`
- `lon_bin`
- `lat_bin`
- `emissions_{pollutant}_mt`

**define broadcasting vs. non-broadcasting?**

## Packages


# Methods 

## Join Emissions Data

Emissions datasets (datasets 1 and 2 above) were read into the pipeline, the column names were converted to snake case, and a new `year-month` column was created for both datasets. In the broadcasting dataset, NA values in the `flag` column were filled with “UNK” to represent flag unknown, and `vessel_class` was filtered for gear types identified with a high degree of confidence (i.e. "squid_jigger", "drifting_longlines", "pole_and_line", "trollers", "pots_and_traps", "set_longlines", "set_gillnets", "trawlers", "dredge_fishing", "tuna_purse_seines", "other_purse_seines", "other_seines"). This eliminated gear types such as "passenger" that were likely mis-identified as "fishing" or as "passenger" by GFW's machine learning algorithm. 

```{r}
unique(emissions$vessel_class)
```

**implications for non-broadcasting**

In the non-broadcasting dataset, emissions estimate columns for each of the 9 pollutants (CO2, CH4, N2O, NOX, SOX, CO, VOCS, PM2.5, PM10) are renamed to match the broadcasting dataset, and a `flag` column is created and populated with “DARK” to distinguish non-broadcasting emissions from the broadcasting emissions. Then, the datasets were concatenated. 

A `year` column was created, and the combined dataset was filtered to 2016-2024 to match the available data for the non-broadcasting dataset. Emissions estimates are then aggregated (summed) by year and flag for each one-by-one degree pixel (distinguished by `lat_bin` and `lon_bin`). 
