---
output: html_document
editor_options: 
  chunk_output_type: console
---

# SAU Validation

This chapter describes how we validated our findings by comparison with Sea Around Us catch data.

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE}
#| label: set-chunk-options
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)

library(tidyverse)

targets::tar_load(full_emissions_fao_species)
targets::tar_load(full_emissions_sau_species)
targets::tar_load(sau_catch)
targets::tar_load(emissions_partitioned_grouped)

source("R/functions.R")
```

## Datasets

[Sea Around Us](https://www.seaaroundus.org/data/#/fao) data was downloaded for each FAO Region (18, 21, 27, 31, 34, 37, 41, 47, 48, 51, 57, 58, 61, 67, 71, 77, 81, 87, and 88).

## Packages

- {tidyverse}
- {janitor}

## Methods

Sea Around US (SAU) data was prepped using the same methods as FAO catch data (see "FAO Data Assembly").

**check that SAU doesn't contain mammals, plants, etc.**
*need to filter for human consumption and industrial*

## Results 

### Overall Comparisons

```{r}
#| code-summary: Overall
# -------- Total emissions comparison --------
# Make sure total emissions match
total_co2 <- sum(emissions_partitioned_grouped$emissions_co2_mt, na.rm = TRUE)
total_co2_fao <- sum(full_emissions_fao_species$total_co2_mt, na.rm = TRUE)
total_co2_sau <- sum(full_emissions_sau_species$total_co2_mt, na.rm = TRUE)

if (total_co2_fao != total_co2_sau) {
  warning("Total emissions in FAO do not match total emissions in SAU.")
}

# -------- Total catch comparison --------
# Filter for years with FAO/SAU overlap (2016-2019)
full_emissions_fao_filtered <- full_emissions_fao_species %>%
  filter(year <= 2019)

full_emissions_sau_filtered <- full_emissions_sau_species %>%
  filter(year <= 2019)

# Compare total catch to determine underreporting
fao_total_catch <- sum(full_emissions_fao_filtered$total_tons_by_species, na.rm = TRUE)
sau_total_catch <- sum(full_emissions_sau_filtered$total_tons_by_species, na.rm = TRUE)

print(paste0("FAO Catch Total: ", round(fao_total_catch, 0), " MT."))
print(paste0("SAU Catch Total: ", round(sau_total_catch, 0), " MT."))
print(paste0("SAU reports ", round(sau_total_catch-fao_total_catch, 0), " more MT of catch."))

# -------- Compare Emissions/Unit Catch --------
fao_total_emissions_catch <- total_co2_fao/fao_total_catch
sau_total_emissions_catch <- total_co2_sau/sau_total_catch

print(paste0("FAO emissions-per-unit-catch: ", round(fao_total_emissions_catch, 2), "."))
print(paste0("SAU emissions-per-unit-catch: ", round(sau_total_emissions_catch, 2), "."))
```

### Comparisons by ISSCAAP Group

```{r}
#| code-summary: ISSCAAP Group
#| fig-height: 10
fao_species <- full_emissions_fao_species %>%
  rename(fao_catch_tons = total_tons_by_species,
         fao_total_co2_mt = total_co2_mt,
        fao_total_ch4_mt = total_ch4_mt,
         fao_total_n2o_mt = total_n2o_mt,
         fao_total_nox_mt = total_nox_mt,
        fao_total_sox_mt = total_sox_mt,
        fao_total_co_mt = total_co_mt,
        fao_total_vocs_mt = total_vocs_mt,
        fao_total_pm2_5_mt = total_pm2_5_mt,
        fao_total_pm10_mt = total_pm10_mt)

sau_species <- full_emissions_sau_species %>%
  rename(sau_catch_tons = total_tons_by_species,
         sau_total_co2_mt = total_co2_mt,
        sau_total_ch4_mt = total_ch4_mt,
         sau_total_n2o_mt = total_n2o_mt,
         sau_total_nox_mt = total_nox_mt,
        sau_total_sox_mt = total_sox_mt,
        sau_total_co_mt = total_co_mt,
        sau_total_vocs_mt = total_vocs_mt,
        sau_total_pm2_5_mt = total_pm2_5_mt,
        sau_total_pm10_mt = total_pm10_mt)

isscaap_comparison <- inner_join(fao_species, sau_species, by = c("zone", "year", "flag", "scientific_name" = "species"))

comp <- isscaap_comparison %>%
  ungroup() %>%
  group_by(isscaap_group, isscaap) %>%
  summarise(fao_co2_emissions = sum(fao_total_co2_mt, na.rm = TRUE),
            fao_catch = sum(fao_catch_tons, na.rm = TRUE),
            sau_co2_emissions = sum(sau_total_co2_mt, na.rm = TRUE),
            sau_catch = sum(sau_catch_tons, na.rm = TRUE)) %>%
  mutate(fao_emissions_unit_catch = fao_co2_emissions/fao_catch,
            sau_emissions_unit_catch = sau_co2_emissions/sau_catch) %>%
  select(isscaap_group, isscaap, fao_emissions_unit_catch, sau_emissions_unit_catch) %>%
  pivot_longer(cols = 3:4, names_to = "entity", values_to = "emissions_unit_catch")

ggplot(comp, aes(reorder(isscaap_group, emissions_unit_catch), emissions_unit_catch, fill = entity)) +
  geom_bar(stat = "identity", position = "dodge") +
   scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  coord_flip() +
  labs(y = "Emissions/Unit Catch",
       x = "ISSCAAP Group",
       fill = "Reporting\nAgency") +
  theme_minimal()
```

