---
title: "Data Pipeline"
author: "Carmen Hoyt"
date: last-modified
format:
  html:
      toc: true
      code-fold: true
editor: visual
execute:
  warning: false
  messages: false
editor_options: 
  chunk_output_type: console
---

# About

This Quarto document hosts the code for the Data Pipeline.

## Load Packages

```{r}
#| code-summary: Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(sf)
```

## Load Emissions Data

```{r}
#| code-summary: Load Emissions Data

# Turn off scientific notation
options(scipen=999)

# Set up import for .csv files in capstone data folder on Workbench-2
file_names <- list.files("/capstone/seamissions/data/meds_capstone_project", pattern="*.csv", full.names=TRUE)

# Read in emissions data files
for (i in seq_along(file_names)) {
  # Detect "non" to name non-broadcasting
  if(str_detect(file_names[i], "non")) {
    table_name <- paste("non_broadcasting")
    assign(table_name, read_csv(file_names[i], show_col_types = FALSE) %>%
             clean_names())
    # Detect "ais" to name broadcasting
  } else if (str_detect(file_names[i], "ais")) {
    table_name <- paste("broadcasting")
    assign(table_name, read_csv(file_names[i], show_col_types = FALSE) %>%
             clean_names())
    # Stop if extra files are detected
  } else {
    stop("Extra file(s) detected.")
  }
}
```

# Merge Emissions Datasets

## Clean Emissions Data

```{r}
#| code-summary: Clean Emissions Data

# Create `year_month` column
broadcasting <- broadcasting %>%
  mutate(date = lubridate::ymd(month)) %>%
  mutate(year_month = format(date, '%Y-%m')) %>%
  # Replace NA values with "UNK"
  mutate(flag = ifelse(is.na(flag), "UNK", flag)) %>%
  # Only select fishing vessels (dataset already only selects for fishing activity)
  filter(vessel_class %in% c("fishing","squid_jigger","drifting_longlines","pole_and_line","other_fishing","trollers","fixed_gear","pots_and_traps","set_longlines","set_gillnets","trawlers","dredge_fishing","seiners","purse_seines","tuna_purse_seines","other_purse_seines","other_seines","driftnets"))
  
non_broadcasting <- non_broadcasting %>%
  mutate(date = lubridate::ymd(month)) %>%
  mutate(year_month = format(date, '%Y-%m')) %>%
  mutate(flag = "DARK")

# Rename pollutant columns in non-broadcasting data to match broadcasting data
colnames(non_broadcasting)[str_detect(colnames(non_broadcasting), "non_broadcasting")] <- str_remove(colnames(non_broadcasting)[str_detect(colnames(non_broadcasting), "non_broadcasting")], "_non_broadcasting") 

# Warning if pollutant column names don't match (for join)
if(unique(colnames(non_broadcasting)[str_detect(colnames(non_broadcasting), "emissions")] == colnames(broadcasting)[str_detect(colnames(broadcasting), "emissions")]) != TRUE) {
  stop("Column names don't match.")
}
```

## Join Emissions Datasets

Join the broadcasting and non-broadcasting datasets.

```{r}
#| code-summary: Join Emissions Datasets

# Combine broadcasting and non-broadcasting data
emissions <- bind_rows(broadcasting, non_broadcasting) %>%
  select(year_month, everything(), -c(month, date))

# Stop if rows don't add up
if(nrow(broadcasting) + nrow(non_broadcasting) != nrow(emissions)) {
  stop("Number of rows don't add up, emissions data lost.")
}
```

### Check for Discrepancies

**This was triggering a very small loss for pm2_5- why?**

```{r}
#| code-summary: Check for Discrepancies

# List pollutants
pollutants <- c("co2", 
                "ch4",
                "n2o",
                "nox",
                "sox",
                "co",
                "vocs",
                "pm2_5",
                "pm10")

# Check for discrepancies
for (i in seq_along(pollutants)) {

  em_col <- paste0("emissions_", pollutants[i], "_mt")
  
  # Total emissions before factoring
  em_before <- sum(broadcasting[[em_col]], na.rm = TRUE) + sum(non_broadcasting[[em_col]], na.rm = TRUE)
  
  # Total emissions after factoring
  em_after <- sum(emissions[[em_col]], na.rm = TRUE)
  
  # Difference in emissions
  diff <- em_after - em_before
  
  # Trigger error warning for lost emissions over 0.1%
  if (diff > 0.001) {
    stop(paste0("Difference in emissions over 0.1% returned for ", {pollutants[i]}, ". Some emissions may be lost."))
  }  
}
```

# Spatial Join FAO Regions and Emissions

## Prepare Emissions Data

```{r}
#| code-summary: Prepare Emissions Data

# -------- EMISSIONS --------
# Filter for desired variables
emissions_filtered <- emissions %>%
  # Create year column
  mutate(year = as.integer(substr(year_month, 1, 4))) %>%
  # Group by pixel, year, and flag
  group_by(lat_bin, lon_bin, year, flag) %>% # vessel_class (?)
  # Sum emissions
  summarise(emissions_co2_mt = sum(emissions_co2_mt, na.rm = TRUE),
            emissions_ch4_mt = sum(emissions_ch4_mt, na.rm = TRUE),
            emissions_n2o_mt = sum(emissions_n2o_mt, na.rm = TRUE),
            emissions_nox_mt = sum(emissions_nox_mt, na.rm = TRUE),
            emissions_sox_mt = sum(emissions_sox_mt, na.rm = TRUE),
            emissions_co_mt = sum(emissions_co_mt, na.rm = TRUE),
            emissions_vocs_mt = sum(emissions_vocs_mt, na.rm = TRUE),
            emissions_pm2_5_mt = sum(emissions_pm2_5_mt, na.rm = TRUE),
            emissions_pm10_mt = sum(emissions_pm10_mt, na.rm = TRUE)) %>%
  # Filter years to match non-broadcasting emissions
  filter(year > 2015)

# Convert lat/long to point geometry (degrees)
emissions_sf <- emissions_filtered %>%
  # add lat and lon bin to preserve for merge 
  dplyr::mutate(lat = lat_bin, lon = lon_bin) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) # import as WGS (unit = degrees)

# Create grid throughout extent of emissions_sf from point geometry
emissions_grid <- emissions_sf %>%
  st_make_grid(cellsize = c(1,1),
               what = "polygons") %>%
  st_sf() %>%
  # Fix invalid geometries
  st_make_valid() %>%
  # Add grid number to keep track
  mutate(grid_id = row_number())
  
# Spatially join emissions data to the grid
emissions_grid_sf <- st_join(emissions_grid, emissions_sf, left = FALSE) %>%
  # Transform to Equal Earth projection
  st_transform(6933)

# Save for dashboard
st_write(emissions_grid_sf, file.path("/capstone/seamissions/checkpoint/emissions_grid_sf.gpkg"), delete_dsn = TRUE)
```

## Load FAO Regions Shapefile and Run Intersection

```{r}
#| code-summary: Run Intersection

# -------- FAO REGIONS --------
# Import regions data (set same crs) from workbench-2
fao_regions <- st_read(file.path("/capstone/seamissions/data/fao_region_shapefile")) %>%
  # Transform to same crs as grid
  st_transform(st_crs(emissions_grid_sf)) %>%
  # Fix geometries
  st_make_valid()

# Confirm crs match
if (st_crs(emissions_grid_sf) != st_crs(fao_regions)){
  stop("CRS don't match.")
}

# Find which grid cells intersect with the FAO regions
emissions_zones <- st_intersection(emissions_grid_sf, fao_regions) 

# Create area column (as numeric)
emissions_zones$area <- as.numeric(st_area(emissions_zones))
```

## Partition Emissions by Proportional Area

```{r}
#| code-summary: Partition emissions 

# Inspect geometry types in emissions_zones
geom_types <- unique(st_geometry_type(emissions_zones))

# Specify desired geometry types: POLYGON or MULTIPOLYGON 
valid_types <- c("POLYGON", "MULTIPOLYGON")

# Identify unexpected types
invalid_types <- setdiff(geom_types, valid_types) # contains POINT (?)

# Warning if geometry types besides POLYGON or MULTIPOLYGON found
if(length(invalid_types > 0)) {
  warning("Additional geometry types detected. Proceed with filter.")
}

# Remove other geometry types (filter out POINT)
emissions_zones_filtered <- emissions_zones %>%
  filter(st_geometry_type(.) %in% valid_types) # in valid_types, c("POLYGON", "MULTIPOLYGON")

# Break MULTIPOLYGONS down into POLYGONS
emissions_zones_exploded <- st_cast(emissions_zones_filtered, "MULTIPOLYGON") %>% 
  st_cast("POLYGON")

# Generate areas by grid_id for sub polygons
emissions_zones_summary_1 <- emissions_zones_exploded %>%
  group_by(grid_id, geometry) %>%
  summarise(
    number_areas = n_distinct(area),
    .groups = "drop") %>% # keep?
  mutate(area_summary = as.numeric(st_area(geometry))) %>%
  ungroup()

# Add up sub polygon areas for overall grid_id area
emissions_zones_summary_2 <- emissions_zones_summary_1 %>%
  group_by(grid_id) %>%
  summarise(
    unique_areas = paste(unique(area_summary), collapse = ", "),
    number_areas = sum(number_areas),
    total_unique_area = sum(area_summary, na.rm = TRUE))

# Made a key of grid_id and overall grid_id area
grid_id_key <- emissions_zones_summary_2 %>%
  select(grid_id, total_unique_area) %>%
  st_drop_geometry()

# Join key with emissions_zones
emissions_zones_joined <- left_join(emissions_zones, grid_id_key)

# build in warning for matching rows? ids?
if(nrow(emissions_zones_joined) != nrow(emissions_zones)) {
  stop("Incorrect join: too many rows.")
}

# Partition out emissions by area proportion
emissions_partitioned <- emissions_zones_joined %>%
  # Calculate proportion
  mutate(prop = area/total_unique_area) %>%
  # Partition emissions
  mutate(prop_co2_mt = emissions_co2_mt * prop,
         prop_ch4_mt = emissions_ch4_mt * prop,
         prop_n2o_mt = emissions_n2o_mt * prop,
         prop_nox_mt = emissions_nox_mt * prop,
         prop_sox_mt = emissions_sox_mt * prop,
         prop_co_mt = emissions_co_mt * prop,
         prop_vocs_mt = emissions_vocs_mt * prop,
         prop_pm2_5_mt = emissions_pm2_5_mt * prop,
         prop_pm10_mt = emissions_pm10_mt * prop)

# Check for discrepancies
for (i in seq_along(pollutants)) {
  
  # Assign column names
  sf_col <- paste0("emissions_", pollutants[i], "_mt")
  prop_col <- paste0("prop_", pollutants[i], "_mt")
  
  # Total emissions before factoring
  em_before <- sum(emissions_grid_sf[[sf_col]], na.rm = TRUE)
  
  # Total emissions after factoring
  em_after <- sum(emissions_partitioned[[prop_col]], na.rm = TRUE)
  
  # Difference in emissions
  diff <- em_after - em_before
  
  # Percent error
  error <- ((diff/em_before)*100)
  
  #print(error)
  
  # Error warning for lost emissions over 0.001% (DECISION POINT)
  if (error > 0.001) {
  stop(paste0("Error over 0.001% returned for ", {pollutants[i]}, " emissions partitioning. Some emissions may be lost."))
  }
}

# Drop geometry
emissions_partitioned_no_geo <- emissions_partitioned %>%
  st_drop_geometry()

# Group by zone, flag, and year
emissions_partitioned_grouped <- emissions_partitioned_no_geo %>%
  group_by(zone, flag, year) %>%
  summarise(emissions_co2_mt = sum(prop_co2_mt, na.rm = TRUE),
            emissions_ch4_mt = sum(prop_ch4_mt, na.rm = TRUE),
            emissions_n2o_mt = sum(prop_n2o_mt, na.rm = TRUE),
            emissions_nox_mt = sum(prop_nox_mt, na.rm = TRUE),
            emissions_sox_mt = sum(prop_sox_mt, na.rm = TRUE),
            emissions_co_mt = sum(prop_co_mt, na.rm = TRUE),
            emissions_vocs_mt = sum(prop_vocs_mt, na.rm = TRUE),
            emissions_pm2_5_mt = sum(prop_pm2_5_mt, na.rm = TRUE),
            emissions_pm10_mt = sum(prop_pm10_mt, na.rm = TRUE)) %>%
  ungroup()

# Save for SAU analysis
write_csv(emissions_partitioned_grouped, file.path("/capstone/seamissions/checkpoint/emissions_partitioned_grouped.csv"))
```

# Join Emissions and Catch Data

## Load FAO Catch Data

```{r}
#| code-summary: Load FAO Catch Data
# Pre-assembled, see data-assembly.qmd
fao_catch <- read_csv(file.path("/capstone/seamissions/checkpoint/fao_cleaned_identifier.csv"), show_col_types = FALSE)
```

## Partition Emissions by Region and Combine

```{r}
#| code-summary: Distribute Emissions

# Define regions
regions <- c(18, 21, 27, 31, 34, 37, 41, 47, 48, 51, 57, 58, 61, 67, 71, 77, 81, 87, 88)

# Create empty vector to append Region #
option_1 <- vector("list", length = length(regions))

# Partition emissions
for (i in seq_along(regions)){
  
 # -------- TABLE 1 --------
  # 1.1 All flags reporting FAO catch in Region #
  total_region_catch <- fao_catch %>%
    filter(zone == regions[i]) %>%
    group_by(zone, year, flag, species_identifier) %>%
    summarise(country_total_tons_by_species = sum(fao_catch_tons, na.rm = TRUE), .groups = "drop") %>%
    ungroup() %>%
    group_by(zone, year) %>%
    mutate(region_total_tons = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
    mutate(prop_fao_catch = country_total_tons_by_species/region_total_tons) %>%
    select(-region_total_tons) %>%
    filter(country_total_tons_by_species > 0)
  
  # 1.2 Summarise non-broadcasting emissions by Region #
  fao_summary_non_broadcasting <- emissions_partitioned_grouped %>%
    # Filter for Region # (to account for emissions later)
    filter(zone == regions[i]) %>%
    # Select non-broadcasting emissions
    filter(flag == "DARK") %>%
    # Rename columns for non-broadcasting emissions
    mutate(non_broad_co2_mt = emissions_co2_mt,
           non_broad_ch4_mt = emissions_ch4_mt,
           non_broad_n2o_mt = emissions_n2o_mt,
           non_broad_nox_mt = emissions_nox_mt,
           non_broad_sox_mt = emissions_sox_mt,
           non_broad_co_mt = emissions_co_mt,
           non_broad_vocs_mt = emissions_vocs_mt,
           non_broad_pm2_5_mt = emissions_pm2_5_mt,
           non_broad_pm10_mt = emissions_pm10_mt) %>%
    # Keep only desired columns
    select(zone, 
           year, 
           non_broad_co2_mt,
           non_broad_ch4_mt,
           non_broad_n2o_mt,
           non_broad_nox_mt,
           non_broad_sox_mt,
           non_broad_co_mt,
           non_broad_vocs_mt,
           non_broad_pm2_5_mt,
           non_broad_pm10_mt)
    # FAO data doesn't go past 2022
    #filter(year <= 2022) 
  
  # TABLE 1: Allocate non-broadcasting emissions to FAO reporting countries by Region #
  total_region_non_broad_allocation <- full_join(total_region_catch, fao_summary_non_broadcasting, by = c("zone", "year")) %>%
    mutate(dist_non_broad_co2_mt = ifelse(is.na(prop_fao_catch), non_broad_co2_mt, prop_fao_catch * non_broad_co2_mt),
           dist_non_broad_ch4_mt = ifelse(is.na(prop_fao_catch), non_broad_ch4_mt, prop_fao_catch * non_broad_ch4_mt),
           dist_non_broad_n2o_mt = ifelse(is.na(prop_fao_catch), non_broad_ch4_mt, prop_fao_catch * non_broad_n2o_mt),
           dist_non_broad_nox_mt = ifelse(is.na(prop_fao_catch), non_broad_nox_mt, prop_fao_catch * non_broad_nox_mt),
           dist_non_broad_sox_mt = ifelse(is.na(prop_fao_catch), non_broad_sox_mt, prop_fao_catch * non_broad_sox_mt),
           dist_non_broad_co_mt = ifelse(is.na(prop_fao_catch), non_broad_co_mt, prop_fao_catch * non_broad_co_mt),
           dist_non_broad_vocs_mt = ifelse(is.na(prop_fao_catch), non_broad_vocs_mt, prop_fao_catch * non_broad_vocs_mt),
           dist_non_broad_pm2_5_mt = ifelse(is.na(prop_fao_catch), non_broad_pm2_5_mt, prop_fao_catch * non_broad_pm2_5_mt),
           dist_non_broad_pm10_mt = ifelse(is.na(prop_fao_catch), non_broad_pm10_mt, prop_fao_catch * non_broad_pm10_mt)
           ) %>%
    #select(-country_total_tons_by_species) %>%
    select(zone,
           year,
           flag,
           species_identifier,
           country_total_tons_by_species,
           dist_non_broad_co2_mt,
           dist_non_broad_ch4_mt,
           dist_non_broad_n2o_mt,
           dist_non_broad_nox_mt,
           dist_non_broad_sox_mt,
           dist_non_broad_co_mt,
           dist_non_broad_vocs_mt,
           dist_non_broad_pm2_5_mt,
           dist_non_broad_pm10_mt) 
  
  # Account for emissions transfer in TABLE 1
  before <- sum(fao_summary_non_broadcasting$non_broad_co2_mt, na.rm = TRUE) 
  after <- sum(total_region_non_broad_allocation$dist_non_broad_co2_mt, na.rm = TRUE) 
  
  percent_diff <- ((after-before)/before)*100
  
  # Warning if emissions loss is greater than 0.001%
  if(percent_diff > 0.001) {
    stop(paste0("Greater than 0.001% of non-broadcasting emissions lost during allocation to FAO reporting countries in Region ", regions[i], ". Check Table 1."))
  }
  
  # -------- TABLE 2 --------
  # 2.1: All flags with broadcasting emissions by Region #
  emissions_partitioned <- emissions_partitioned_grouped %>%
    # Filter for Region #
    filter(zone == regions[i]) %>%
    # Remove non-broadcasting emissions
    filter(!flag == "DARK") %>%
    mutate(broad_co2_mt = emissions_co2_mt,
           broad_ch4_mt = emissions_ch4_mt,
           broad_n2o_mt = emissions_n2o_mt,
           broad_nox_mt = emissions_nox_mt,
           broad_sox_mt = emissions_sox_mt,
           broad_co_mt = emissions_co_mt,
           broad_vocs_mt = emissions_vocs_mt,
           broad_pm2_5_mt = emissions_pm2_5_mt,
           broad_pm10_mt = emissions_pm10_mt) %>%
    select(-c(emissions_co2_mt, emissions_ch4_mt, emissions_n2o_mt, emissions_nox_mt, emissions_sox_mt, emissions_co_mt, emissions_vocs_mt, emissions_pm2_5_mt, emissions_pm10_mt))
  
  # 2.2 Distribute broadcasting for countries with catch ASSUMPTION: 1 ton = same emissions for all species
  total_flag_catch <- fao_catch %>%
    filter(zone == regions[i]) %>%
    group_by(zone, year, flag, species_identifier) %>%
    summarise(country_total_tons_by_species = sum(fao_catch_tons, na.rm = TRUE), .groups = "drop") %>%
    ungroup() %>%
    group_by(zone, year, flag) %>%
    mutate(country_total_tons = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
    mutate(prop_species_catch = country_total_tons_by_species/country_total_tons) %>%
    filter(country_total_tons_by_species > 0) %>%
    select(-c(country_total_tons)) %>% #country_total_tons_by_species
    filter(!is.na(prop_species_catch))
  
  # TABLE 2
  total_region_broad_allocation <- full_join(total_flag_catch, emissions_partitioned, by = c("zone", "year", "flag")) %>%
    mutate(dist_broad_co2_mt = ifelse(is.na(prop_species_catch), broad_co2_mt, prop_species_catch * broad_co2_mt),
           dist_broad_ch4_mt = ifelse(is.na(prop_species_catch), broad_ch4_mt, prop_species_catch * broad_ch4_mt),
           dist_broad_n2o_mt = ifelse(is.na(prop_species_catch), broad_n2o_mt, prop_species_catch * broad_n2o_mt),
           dist_broad_nox_mt = ifelse(is.na(prop_species_catch), broad_nox_mt, prop_species_catch * broad_nox_mt),
           dist_broad_sox_mt = ifelse(is.na(prop_species_catch), broad_sox_mt, prop_species_catch * broad_sox_mt),
           dist_broad_co_mt = ifelse(is.na(prop_species_catch), broad_co_mt, prop_species_catch * broad_co_mt),
           dist_broad_vocs_mt = ifelse(is.na(prop_species_catch), broad_vocs_mt, prop_species_catch * broad_vocs_mt),
           dist_broad_pm2_5_mt = ifelse(is.na(prop_species_catch), broad_pm2_5_mt, prop_species_catch * broad_pm2_5_mt),
           dist_broad_pm10_mt = ifelse(is.na(prop_species_catch), broad_pm10_mt, prop_species_catch * broad_pm10_mt)) %>%
    select(-country_total_tons_by_species)
  
  # fill taxonomic code NA with "non-catch associated emissions"
  
  # -------- TABLE 3 --------
  # TABLE 3: Full join Tables 1 and 2 (to not lose flags with non-broad emissions but no broad emissions)
  total_region_emissions <- full_join(total_region_broad_allocation, total_region_non_broad_allocation, by = c("zone", "year", "flag", "species_identifier")) %>%
    #filter(zone == 18) %>%
    # Sum broadcasting and non-broadcasting emissions
    mutate(total_co2_mt = rowSums(across(c(dist_broad_co2_mt, dist_non_broad_co2_mt)), na.rm = TRUE),
           total_ch4_mt = rowSums(across(c(dist_broad_ch4_mt, dist_non_broad_ch4_mt)), na.rm = TRUE),
           total_n2o_mt = rowSums(across(c(dist_broad_n2o_mt, dist_non_broad_n2o_mt)), na.rm = TRUE),
           total_nox_mt = rowSums(across(c(dist_broad_nox_mt, dist_non_broad_nox_mt)), na.rm = TRUE),
           total_sox_mt = rowSums(across(c(dist_broad_sox_mt, dist_non_broad_sox_mt)), na.rm = TRUE),
           total_co_mt = rowSums(across(c(dist_broad_co_mt, dist_non_broad_co_mt)), na.rm = TRUE),
           total_vocs_mt = rowSums(across(c(dist_broad_vocs_mt, dist_non_broad_vocs_mt)), na.rm = TRUE),
           total_pm2_5_mt = rowSums(across(c(dist_broad_pm2_5_mt, dist_non_broad_pm2_5_mt)), na.rm = TRUE),
           total_pm10_mt = rowSums(across(c(dist_broad_pm10_mt, dist_non_broad_pm10_mt)), na.rm = TRUE)) %>%
  mutate(species_identifier = ifelse(is.na(species_identifier), 99999999, species_identifier))
  
  # Keep desired columns
  total_region_emissions <- total_region_emissions %>%
    select(zone,
           year,
           flag,
           species_identifier,
           country_total_tons_by_species,
           total_co2_mt,
           total_ch4_mt,
           total_n2o_mt,
           total_nox_mt,
           total_sox_mt,
           total_co_mt,
           total_pm2_5_mt,
           total_pm10_mt
    )
  
  # BUILD SPECIES TABLE for zone, year, flag, species, catch_tons
  
  # Account for emissions after full join
  before <- sum(emissions_partitioned$broad_co2_mt, na.rm = TRUE) + sum(fao_summary_non_broadcasting$non_broad_co2_mt, na.rm = TRUE)
  after <- sum(total_region_emissions$total_co2_mt, na.rm = TRUE) 
  
  percent_diff <- ((after-before)/before)*100
  
  # Warning if emissions loss is greater than 0.001%
  if(percent_diff > 0.001) {
    stop(paste0("Greater than 0.001% of emissions lost during full join of broadcasting and non-broadcasting in Region ", regions[i], ". Check Table 3."))
  }
  
  check <- emissions_partitioned_grouped %>%
    # Filter for Region #
    filter(zone == regions[i])

  before <- sum(check$emissions_co2_mt, na.rm = TRUE)
  after <- sum(total_region_emissions$total_co2_mt, na.rm = TRUE)
  
  percent_diff <- ((after-before)/before)*100
  
  # Warning if emissions loss is greater than 0.001%
  if(percent_diff > 0.001) {
    stop(paste0("Greater than 0.001% of emissions lost during full partitioning in Region ", regions[i], "."))
  }
  
  # Append to vector
  option_1[[i]] <- total_region_emissions
}

# Full emissions dataset
full_emissions_fao <- bind_rows(option_1)

# WARNING TO DETECT ALL REGIONS?

write_csv(full_emissions_fao, file.path("/capstone/seamissions/checkpoint/full_emissions_fao.csv"))
```

## Final Check

```{r}
#| code-summary: Final Check
# Make sure total emissions match
total_co2 <- sum(emissions_partitioned_grouped$emissions_co2_mt, na.rm = TRUE)
total_co2_fao <- sum(full_emissions_fao$total_co2_mt, na.rm = TRUE)

percent_diff <- ((total_co2_fao-total_co2)/total_co2)*100
  
if(percent_diff > 0.001) {
  warning("Emissions lost in FAO proportioning.")
}
```

