---
title: "Emissions Pipeline"
date: last-modified
format:
  html:
      toc: true
      code-fold: true
editor: visual
execute:
  warning: false
  messages: false
editor_options: 
  chunk_output_type: console
---

# About

This repository hosts the code for the open-source data processing pipeline.

## Load packages

```{r}
#| code-summary: Expand code
library(tidyverse)
library(janitor)
library(here)
#library(gghighlight)
library(lubridate)
#library(ggwordcloud)
#library(glue)
library(sf)
library(dplyr)
```

## Load data

```{r}
#| code-summary: Load project data

# Turn off scientific notation
options(scipen=999)

# ---- Load project data ----
data_directory_base <-  ifelse(Sys.info()["nodename"] == "quebracho" | Sys.info()["nodename"] == "sequoia",
                               "/home/emlab",
                               # Otherwise, set the directory for local machines based on the OS
                               # If using Mac OS, the directory will be automatically set as follows
                               ifelse(Sys.info()["sysname"]=="Darwin",
                                      "/Users/Shared/nextcloud/emLab",
                                      # If using Windows, the directory will be automatically set as follows
                                      ifelse(Sys.info()["sysname"]=="Windows",
                                             "G:/Shared\ drives/nextcloud/emLab",
                                             # If using Linux, will need to manually modify the following directory path based on their user name
                                             # Replace your_username with your local machine user name
                                             "/home/choyt/Nextcloud")))

project_directory <- paste0(data_directory_base, "/projects/current-projects/ocean-ghg-fisheries/data/processed/gfw-emissions/")
```

```{r}
#| code-summary: Load emissions data

# Turn off scientific notation
options(scipen=999)

# Set file path to data folder workbench-2
pathway <- "/capstone/seamissions/data/meds_capstone_project"

# Set up import for .csv or .xlsx files
file_names <- list.files(pathway, pattern="*.csv", full.names=TRUE)

# Read in emissions data files
for (i in seq_along(file_names)) {
 if(str_detect(file_names[i], "non")) {
    table_name <- paste("non_broadcasting")
    assign(table_name, read_csv(file_names[i], show_col_types = FALSE) %>%
          clean_names())

  } else if (str_detect(file_names[i], "ais")) {
    table_name <- paste("broadcasting")
    assign(table_name, read_csv(file_names[i], show_col_types = FALSE) %>%
          clean_names())
    
  } else {
    warning("Extra file detected.")
  }
}
```

# Merge Emissions Datasets

## Clean Data

```{r}
#| code-summary: Clean Data

# Create `year_month` column
broadcasting <- broadcasting %>%
  mutate(date = lubridate::ymd(month)) %>%
  mutate(year_month = format(date, '%Y-%m'))
  
# Create `year_month` column 
non_broadcasting <- non_broadcasting %>%
  mutate(date = lubridate::ymd(month)) %>%
  mutate(year_month = format(date, '%Y-%m'))
```

## Join Data

```{r}
#| code-summary: Join Broadcasting and Non-Broadcasting Datasets

# Left join broadcasting and non-broadcasting datasets
emissions_joined <- left_join(broadcasting, non_broadcasting, by = c("year_month", "lat_bin", "lon_bin", "length_size_class_percentile")) %>%
  
  # Replace non-broadcasting NA values with (see decision point below)
  mutate(emissions_co2_non_broadcasting_mt = ifelse(is.na(emissions_co2_non_broadcasting_mt), 0, emissions_co2_non_broadcasting_mt),
         emissions_ch4_non_broadcasting_mt = ifelse(is.na(emissions_ch4_non_broadcasting_mt), 0, emissions_ch4_non_broadcasting_mt),
         emissions_n2o_non_broadcasting_mt = ifelse(is.na(emissions_n2o_non_broadcasting_mt), 0, emissions_n2o_non_broadcasting_mt),
         emissions_nox_non_broadcasting_mt = ifelse(is.na(emissions_nox_non_broadcasting_mt), 0, emissions_nox_non_broadcasting_mt),
         emissions_sox_non_broadcasting_mt = ifelse(is.na(emissions_sox_non_broadcasting_mt), 0, emissions_sox_non_broadcasting_mt),
         emissions_pm_non_broadcasting_mt = ifelse(is.na(emissions_pm_non_broadcasting_mt), 0, emissions_pm_non_broadcasting_mt),
         emissions_co_non_broadcasting_mt = ifelse(is.na(emissions_co_non_broadcasting_mt), 0, emissions_co_non_broadcasting_mt),
         emissions_vocs_non_broadcasting_mt = ifelse(is.na(emissions_vocs_non_broadcasting_mt), 0, emissions_co_non_broadcasting_mt),
         emissions_pm2_5_non_broadcasting_mt = ifelse(is.na(emissions_pm2_5_non_broadcasting_mt), 0, emissions_co_non_broadcasting_mt),
         emissions_pm10_non_broadcasting_mt = ifelse(is.na(emissions_pm10_non_broadcasting_mt), 0, emissions_co_non_broadcasting_mt)) %>%

  # Select columns to keep
  dplyr::select(year_month, 
         lon_bin,
         lat_bin,
         flag, 
         vessel_class, 
         length_size_class_percentile, 
         fishing_activity, 
         hours, 
         kw_hours,
         emissions_co2_mt, 
         emissions_ch4_mt, 
         emissions_n2o_mt, 
         emissions_nox_mt, 
         emissions_sox_mt, 
         emissions_co_mt,
         emissions_vocs_mt,
         emissions_pm2_5_mt,
         emissions_pm10_mt,
         emissions_co2_non_broadcasting_mt,
         emissions_ch4_non_broadcasting_mt,
         emissions_n2o_non_broadcasting_mt,
         emissions_nox_non_broadcasting_mt,
         emissions_sox_non_broadcasting_mt,
         emissions_pm_non_broadcasting_mt,
         emissions_co_non_broadcasting_mt,
         emissions_vocs_non_broadcasting_mt,
         emissions_pm2_5_non_broadcasting_mt,
         emissions_pm10_non_broadcasting_mt
         )
```

**Decision point:** NA values in `emissions_non_broadcasting` are giving NA in the totals after factoring (`non_broadcasting` missing 2015 data, NA values on join).

-   Solution: replaced NA with 0
-   Filter `broadcasting` for year \>= 2016?

## Factor Emissions

```{r}
#| code-summary: Expand code

# Factor non-broadcasting emissions
emissions <- emissions_joined %>%
  
  # Calculate total ais emissions per grouped effort
  group_by(year_month, lon_bin, lat_bin, flag, vessel_class, length_size_class_percentile, hours, kw_hours) %>% 
  mutate(ais_emissions_co2 = sum(emissions_co2_mt, na.rm = TRUE),
         ais_emissions_ch4 = sum(emissions_ch4_mt, na.rm = TRUE),
         ais_emissions_n2o = sum(emissions_n2o_mt, na.rm = TRUE),
         ais_emissions_nox = sum(emissions_nox_mt, na.rm = TRUE),
         ais_emissions_sox = sum(emissions_sox_mt, na.rm = TRUE),
         ais_emissions_co = sum(emissions_co_mt, na.rm = TRUE),
         ais_emissions_vocs = sum(emissions_vocs_mt, na.rm = TRUE),
         ais_emissions_pm2_5 = sum(emissions_pm2_5_mt, na.rm = TRUE),
         ais_emissions_pm10 = sum(emissions_pm10_mt, na.rm = TRUE)
         ) %>%
  
  # Calculate total ais emissions per pixel and length size class
  group_by(year_month, lon_bin, lat_bin, length_size_class_percentile) %>%
  mutate(total_ais_emissions_co2 = sum(emissions_co2_mt, na.rm = TRUE),
         total_ais_emissions_ch4 = sum(emissions_ch4_mt, na.rm = TRUE),
         total_ais_emissions_n2o = sum(emissions_n2o_mt, na.rm = TRUE),
         total_ais_emissions_nox = sum(emissions_nox_mt, na.rm = TRUE),
         total_ais_emissions_sox = sum(emissions_sox_mt, na.rm = TRUE),
         total_ais_emissions_co = sum(emissions_co_mt, na.rm = TRUE),
         total_ais_emissions_vocs = sum(emissions_vocs_mt, na.rm = TRUE),
         total_ais_emissions_pm2_5 = sum(emissions_pm2_5_mt, na.rm = TRUE),
         total_ais_emissions_pm10 = sum(emissions_pm10_mt, na.rm = TRUE)) %>%
  ungroup() %>%
  
  # Define a factor for the proportion of hours per flag/total hours 
  mutate(factor_co2 = ais_emissions_co2/total_ais_emissions_co2,
         factor_ch4 = ais_emissions_ch4/total_ais_emissions_ch4,
         factor_n2o = ais_emissions_n2o/total_ais_emissions_n2o,
         factor_nox = ais_emissions_nox/total_ais_emissions_nox,
         factor_sox = ais_emissions_sox/total_ais_emissions_sox,
         factor_co = ais_emissions_co/total_ais_emissions_co,
         factor_vocs = ais_emissions_vocs/total_ais_emissions_vocs,
         factor_pm2_5 = ais_emissions_pm2_5/total_ais_emissions_pm2_5,
         factor_pm10 = ais_emissions_pm10/total_ais_emissions_pm10) %>%
  
  # Apply factor to emissions columns to partition nb data
  mutate(total_co2 = emissions_co2_mt + (factor_co2 * emissions_co2_non_broadcasting_mt),
         total_ch4 = emissions_ch4_mt + (factor_ch4 * emissions_ch4_non_broadcasting_mt),
         total_n2o = emissions_n2o_mt + (factor_n2o * emissions_n2o_non_broadcasting_mt),
         total_nox = emissions_nox_mt + (factor_nox * emissions_nox_non_broadcasting_mt),
         total_sox = emissions_sox_mt + (factor_sox * emissions_sox_non_broadcasting_mt),
         total_co = emissions_co_mt + (factor_co * emissions_co_non_broadcasting_mt),
         total_vocs = emissions_vocs_mt + (factor_vocs * emissions_vocs_non_broadcasting_mt),
         total_pm2_5 =  emissions_pm2_5_mt + (factor_pm2_5 * emissions_pm2_5_non_broadcasting_mt),
         total_pm10 = emissions_pm10_mt + (factor_pm10 * emissions_pm10_non_broadcasting_mt)
         ) %>%

  # Select columns of interest
  dplyr::select(year_month, 
         lon_bin,
         lat_bin,
         flag, 
         vessel_class, 
         length_size_class_percentile, 
         fishing_activity, 
         hours, 
         kw_hours,
         total_co2, 
         total_ch4,
         total_n2o,
         total_nox,
         total_sox,
         total_co,
         total_vocs,
         total_pm2_5,
         total_pm10
         )
```

## Check for Discrepancies

```{r}
#| code-summary: Expand code

# List pollutants
pollutants <- c("co2", 
                "ch4",
                "n2o",
                "nox",
                "sox",
                "co",
                "vocs",
                "pm2_5",
                "pm10")

# Check for discrepancies
for (i in seq_along(pollutants)) {
  
  # Assign column names
  broad_col <- paste0("emissions_", pollutants[i], "_mt")
  non_broad_col <- paste0("emissions_", pollutants[i], "_non_broadcasting_mt")
  em_col <- paste0("total_", pollutants[i])
  
  # Total emissions before factoring
  em_before <- sum(broadcasting[[broad_col]], na.rm = TRUE) + sum(non_broadcasting[[non_broad_col]], na.rm = TRUE)
  
  # Total emissions after factoring
  em_after <- sum(emissions[[em_col]], na.rm = TRUE)
  
  # Difference in emissions
  diff <- em_after - em_before
  
  # Percent error
  error <- diff/em_before
  
  # Trigger error warning for lost emissions over 0.1%
  if (error > 0.001) {
  warning(paste0("Error over 0.1% returned for ", {pollutants[i]}, " emissions partitioning. Some emissions may be lost."))
  }
}
```

## Pivot Data for Visualizations

```{r}
#| code-summary: Expand code

# Pivot emissions data
em_pivot <- emissions %>%
  pivot_longer(10:18,
                names_to = "pollutant_type",
                values_to = "emissions") %>%
  mutate(pollutant_type = str_remove(pollutant_type, "total_"),
         year = as.numeric(substring(year_month, 1, 4)))
```

### Checkpoint

```{r}
#| code-summary: Expand code

# -------- Create "checkpoint" folder if needed --------

# Save emissions dataset to workbench-2
write_csv(emissions, file.path("/capstone/seamissions/checkpoint/emissions.csv"))

# Read in emissions dataset
#emissions <- read_csv(file.path("/capstone/seamissions/checkpoint/emissions.csv"), show_col_types = FALSE)

# Save pivot dataset to workbench-2
write_csv(em_pivot, file.path("/capstone/seamissions/checkpoint/em_pivot.csv"))

# Read in pivot dataset to workbench-2
#em_pivot <- read_csv(file.path("/capstone/seamissions/checkpoint/em_pivot.csv"), show_col_types = FALSE)
```

# Spatial Join FAO Regions and Emissions

## Import and clean data

```{r}
#| code-summary: Expand code

# -------- set up projections --------
# equal earth 8857
#my_crs <- st_crs("+proj=eqearth +datum=WGS84 +units=m +no_defs")

# OR mollweide projection 54009 (but not recognized)
#mollweide_proj <- "+proj=moll +lon_0=0 +datum=WGS84 +units=m +no_defs"
#emissions_sf <- st_set_crs(emissions_sf, mollweide_proj) # transform?

# -------- EMISSIONS --------
# Filter for desired variables
emissions_filtered <- emissions %>%
  # Create year column
  mutate(year = as.integer(substr(year_month, 1, 4))) %>%
  # Group by pixel, year, and flag
  group_by(lat_bin, lon_bin, year, flag) %>% # vessel_class
  # Sum co2 (ADD OTHER POLLUTANTS HERE)
  summarise(total_co2_mt = sum(total_co2, na.rm = TRUE))

# Convert lat/long to point geometry (degrees)
emissions_sf <- emissions_filtered %>%
  # add lat and lon bin to preserve for merge 
  dplyr::mutate(lat = lat_bin, lon = lon_bin) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) # import as WGS (unit = degrees)

# Create grid throughout extent of emissions_sf from point geometry
emissions_grid <- emissions_sf %>%
  st_make_grid(cellsize = c(1,1),
               what = "polygons") %>%
  st_sf() %>%
  st_make_valid()
  
# Spatially join emissions data to the grid
emissions_grid_sf <- st_join(emissions_grid, emissions_sf, left = FALSE) %>%
  # Transform to Equal Earth projection
  st_transform(6933)

# Add area column before intersection
#emissions_grid_sf$area_before <- st_area(emissions_grid_sf)

# -------- FAO REGIONS --------
# Import regions data (set same crs) from workbench-2
fao_regions <- st_read(file.path("/capstone/seamissions/data/fao_region_shapefile")) %>%
  # Transform to same crs as grid
  st_transform(st_crs(emissions_grid_sf)) %>%
  # Fix geometries
  st_make_valid()

# Confirm crs match
if (st_crs(emissions_grid_sf) != st_crs(fao_regions)){
  warning("CRS don't match.")
}

# -------- FAO CATCH --------
# Import catch data from workbench-2
fao_catch <- read_csv(file.path("/capstone/seamissions/data/fao_seafood_production/fao_catch.csv"), show_col_types = FALSE) %>%
  # Select desired columns
  select(area_code,
         period,
         iso3_code,
         official_name_en,
         scientific_name,
         value) %>%
  # Rename for consistency
  rename(zone = area_code,
         year = period,
         flag = iso3_code,
         catch = value,
         official_name = official_name_en) %>%
  # Filter for >2015 to match emissions data
  filter(year >= 2015) %>%
  group_by(year, zone, flag, scientific_name, official_name) %>%
  summarise(total_catch_t = sum(catch, na.rm = TRUE)) %>%
  mutate(zone = as.integer(zone),
         year = as.integer(year))

```

### Checkpoint

```{r}
#| code-summary: Checkpoint

# Save emissions_grid_sf to workbench-2
#st_write(emissions_grid_sf, file.path("/capstone/seamissions/checkpoint/emissions_grid_sf.gpkg"), delete_dsn = TRUE)

# Read in emissions_grid_sf from workbench-2
emissions_grid_sf <- st_read(file.path("/capstone/seamissions/checkpoint/emissions_grid_sf.gpkg"))

# Save cleaned fao_catch data
#write_csv(fao_catch, file.path("/capstone/seamissions/checkpoint/fao_catch.csv"))

# Read in fao catch data
fao_catch <- read_csv(file.path("/capstone/seamissions/checkpoint/fao_catch.csv"), show_col_types = FALSE)
```

## Run intersects

```{r}
# Assign majority pixel to an FAO region 
pixels_with_fao_region <- st_join(emissions_grid_sf, fao_regions, largest = TRUE)

# Checkpoint
st_write(pixels_with_fao_region, file.path("/capstone/seamissions/checkpoint/pixels_with_fao_region.gpkg"), delete_dsn = TRUE)

# check
if(nrow(pixels_with_fao_region) != nrow(emissions_grid_sf)) {
  warning("Number of rows differs.")
} 

# check
zone_na <- pixels_with_fao_region %>%
  filter(is.na(zone))
zone_31 <- pixels_with_fao_region %>%
  filter(zone == 31)
fao_zone_31 <- fao_regions %>%
  filter(zone == 31)

plot(st_geometry(zone_31))
plot(st_geometry(fao_zone_31), add = TRUE)
```

```{r}
#| code-summary: Checkpoint
# Find which grid cells intersect with the FAO regions
overlapping_pixels <- st_intersects(emissions_grid_sf, fao_regions) 

# Select grid polygons that have an intersection (TRUE values)
emissions_zones <- emissions_grid_sf[lengths(overlapping_pixels) > 0, ]

# Add regions column
emissions_zones$region <- sapply(overlapping_pixels[lengths(overlapping_pixels) > 0], function(x) {
  if (length(x) > 0) {
    return(paste(fao_regions$zone[x], collapse = ","))
  } else {
    return(NA)
  }
})

# Filter grid cells that intersect with more than one FAO region
multiple_regions <- emissions_zones[grepl(",", emissions_zones$region), ]

# Drop geometry
emissions_zones_no_geo <- emissions_zones %>%
  st_drop_geometry()

#overlapping_grids <- st_difference(multiple_regions, pixels_with_fao_region)  
common_grids <- st_intersection(multiple_regions, pixels_with_fao_region) # Cells that appear in both
all_grids <- st_union(multiple_regions, pixels_with_fao_region)           # All grid cells
unique_grids <- st_difference(all_grids, common_grids) # keep only unique grid cells
```

### Checkpoint

```{r}
#| code-summary: Checkpoint

# Save emissions_zones to workbench-2
#st_write(emissions_zones, file.path("/capstone/seamissions/checkpoint/emissions_zones.gpkg"), delete_dsn = TRUE)

# Read in emissions_zones from workbench-2
emissions_zones <- st_read(file.path("/capstone/seamissions/checkpoint/emissions_zones.gpkg"))

# Save cleaned emissions_zones_no_geo data
write_csv(emissions_zones_no_geo, file.path("/capstone/seamissions/checkpoint/emissions_zones_no_geo.csv"))

# Read in emissions_zones_no_geo data
#emissions_zones_no_geo <- read_csv(file.path("/capstone/seamissions/checkpoint/emissions_zones_no_geo.csv"), show_col_types = FALSE)
```
