---
title: "FAO-SAU Comparison"
author: "Carmen Hoyt"
date: last-modified
format:
  html:
      toc: true
      code-fold: true
editor: visual
execute:
  warning: false
  messages: false
editor_options: 
  chunk_output_type: console
---

# About

This document compares FAO emissions-per-unit-catch results to SAU emissions-per-unit-catch results.

# Prepare SAU data

## Methods:

-   For each year, calculate total **non-broadcasting** emissions for SAU data for each FAO region: `sau_summary_non_broadcasting`
-   Then divy that up to *all* fisheries that report catch in that region (proportionally by the amount of catch in each fishery). TABLE 1: `total_region_non_broad_allocation`
-   For each year, calculate total **broadcasting** emissions, by country, for SAU data for each FAO region. TABLE 2: `emissions_partitioned`
-   Full join the allocated non-broadcasting emissions with the broadcasting emissions. TABLE 3: `total_region_emissions`

Assumptions:

-   Within an FAO region, the non-broadcasting emissions are spread proportionally across *all* fisheries and countries within that region *by catch* in the SAU dataset.

### Load Packages

```{r}
#| code-summary: Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(sf)
library(gghighlight)
```

### Load Data

```{r}
#| code-summary: Load Data
# Read in emissions data
emissions_partitioned_grouped <- read_csv(file.path("/capstone/seamissions/checkpoint/emissions_partitioned_grouped.csv"), show_col_types = FALSE)

# Read in FAO country code table
# iso_flags <- read_csv("/capstone/seamissions/data/fao_seafood_production/CL_FI_COUNTRY_GROUPS.csv", show_col_types = FALSE) %>%
#   clean_names() %>%
#   select(iso3_code, name_en) 

# Read in flag key
flag_key <- read_csv(file.path("/capstone/seamissions/checkpoint/flag_key.csv"), show_col_types = FALSE)

# Set up import for .csv files
file_names <- list.files("/capstone/seamissions/data/sau", pattern="*.csv", full.names = TRUE)

sau_regions <- vector("list", length = length(file_names))

# Read in SAU files by region
for (i in seq_along(file_names)) {
  table_name <- paste0("sau_region_", str_sub(file_names[i], 40, 41))
  sau_regions[[i]] <- assign(table_name, read.csv(file_names[i], stringsAsFactors = FALSE, check.names = TRUE) %>%
                               clean_names() %>%
                               mutate(zone = as.numeric(str_sub(file_names[i], 40, 41)),
                                      year = as.numeric(year)) %>%
                               filter(year >= 2015) %>%
                               # WORK ON THIS
                               left_join(flag_key, by = c("fishing_entity" = "sau_name")) %>%
                               rename(flag = iso3_code)
                               )
}

# Bind all regions
sau_catch_data <- bind_rows(sau_regions)

# Summarise SAU catch
sau_catch_summary <- sau_catch_data %>%
  group_by(zone, year, flag, scientific_name, common_name, commercial_group) %>%
  summarise(sau_catch_tons = sum(tonnes, na.rm = TRUE)) %>%
  rename(species = scientific_name) %>% 
  filter(year > 2015)
```

### Partition Emissions by Region and Combine

```{r}
#| code-summary: Distribute Emissions

# Turn off scientific notation
options(scipen = 999)

# Define regions
regions <- c(18, 21, 27, 31, 34, 37, 41, 47, 48, 51, 57, 58, 61, 67, 71, 77, 81, 87, 88)

# Create empty vector to append Region #
option_1 <- vector("list", length = length(regions))

# Partition emissions
for (i in seq_along(regions)){

  # -------- TABLE 1 --------
  # 1.1 All flags reporting FAO catch in Region #
  total_region_catch <- sau_catch_summary %>%
    filter(zone == regions[i]) %>%
    group_by(zone, year, flag, species) %>%
    summarise(country_total_tons_by_species = sum(sau_catch_tons, na.rm = TRUE), .groups = "drop") %>%
    ungroup() %>%
    group_by(zone, year) %>%
    mutate(region_total_tons = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
    mutate(prop_sau_catch = country_total_tons_by_species/region_total_tons) %>%
    select(-region_total_tons) %>%
    filter(country_total_tons_by_species > 0)
  
  # 1.2 Summarise non-broadcasting emissions by Region #
  sau_summary_non_broadcasting <- emissions_partitioned_grouped %>%
    # Filter for Region # (to account for emissions later)
    filter(zone == regions[i]) %>%
    # Select non-broadcasting emissions
    filter(flag == "DARK") %>%
    # Rename columns for non-broadcasting emissions
    mutate(non_broad_co2_mt = emissions_co2_mt,
           non_broad_ch4_mt = emissions_ch4_mt,
           non_broad_n2o_mt = emissions_n2o_mt,
           non_broad_nox_mt = emissions_nox_mt,
           non_broad_sox_mt = emissions_sox_mt,
           non_broad_co_mt = emissions_co_mt,
           non_broad_vocs_mt = emissions_vocs_mt,
           non_broad_pm2_5_mt = emissions_pm2_5_mt,
           non_broad_pm10_mt = emissions_pm10_mt) %>%
    # Keep only desired columns
    select(zone, 
           year, 
           non_broad_co2_mt,
           non_broad_ch4_mt,
           non_broad_n2o_mt,
           non_broad_nox_mt,
           non_broad_sox_mt,
           non_broad_co_mt,
           non_broad_vocs_mt,
           non_broad_pm2_5_mt,
           non_broad_pm10_mt)
  
  # TABLE 1: Allocate non-broadcasting emissions to FAO reporting countries by Region #
  total_region_non_broad_allocation <- full_join(total_region_catch, sau_summary_non_broadcasting, by = c("zone", "year")) %>%
    mutate(dist_non_broad_co2_mt = ifelse(is.na(prop_sau_catch), non_broad_co2_mt, prop_sau_catch * non_broad_co2_mt),
           dist_non_broad_ch4_mt = ifelse(is.na(prop_sau_catch), non_broad_ch4_mt, prop_sau_catch * non_broad_ch4_mt),
           dist_non_broad_n2o_mt = ifelse(is.na(prop_sau_catch), non_broad_ch4_mt, prop_sau_catch * non_broad_n2o_mt),
           dist_non_broad_nox_mt = ifelse(is.na(prop_sau_catch), non_broad_nox_mt, prop_sau_catch * non_broad_nox_mt),
           dist_non_broad_sox_mt = ifelse(is.na(prop_sau_catch), non_broad_sox_mt, prop_sau_catch * non_broad_sox_mt),
           dist_non_broad_co_mt = ifelse(is.na(prop_sau_catch), non_broad_co_mt, prop_sau_catch * non_broad_co_mt),
           dist_non_broad_vocs_mt = ifelse(is.na(prop_sau_catch), non_broad_vocs_mt, prop_sau_catch * non_broad_vocs_mt),
           dist_non_broad_pm2_5_mt = ifelse(is.na(prop_sau_catch), non_broad_pm2_5_mt, prop_sau_catch * non_broad_pm2_5_mt),
           dist_non_broad_pm10_mt = ifelse(is.na(prop_sau_catch), non_broad_pm10_mt, prop_sau_catch * non_broad_pm10_mt)
           ) %>%
    select(zone,
           year,
           flag,
           species,
           country_total_tons_by_species,
           dist_non_broad_co2_mt,
           dist_non_broad_ch4_mt,
           dist_non_broad_n2o_mt,
           dist_non_broad_nox_mt,
           dist_non_broad_sox_mt,
           dist_non_broad_co_mt,
           dist_non_broad_vocs_mt,
           dist_non_broad_pm2_5_mt,
           dist_non_broad_pm10_mt) 
  
  # Account for emissions transfer in TABLE 1
  before <- sum(sau_summary_non_broadcasting$non_broad_co2_mt, na.rm = TRUE) 
  after <- sum(total_region_non_broad_allocation$dist_non_broad_co2_mt, na.rm = TRUE) 
  
  percent_diff <- ((after-before)/before)*100
  
  # Warning if emissions loss is greater than 0.001%
  if(percent_diff > 0.001) {
    warning(paste0("Greater than 0.001% of non-broadcasting emissions lost during allocation to SAU reporting countries in Region ", regions[i], ". Check Table 1."))
  }
  
  # -------- TABLE 2 --------
  # 2.1: All flags with broadcasting emissions by Region #
  emissions_partitioned <- emissions_partitioned_grouped %>%
    # Filter for Region #
    filter(zone == regions[i]) %>%
    # Remove non-broadcasting emissions
    filter(!flag == "DARK") %>%
    mutate(broad_co2_mt = emissions_co2_mt,
           broad_ch4_mt = emissions_ch4_mt,
           broad_n2o_mt = emissions_n2o_mt,
           broad_nox_mt = emissions_nox_mt,
           broad_sox_mt = emissions_sox_mt,
           broad_co_mt = emissions_co_mt,
           broad_vocs_mt = emissions_vocs_mt,
           broad_pm2_5_mt = emissions_pm2_5_mt,
           broad_pm10_mt = emissions_pm10_mt) %>%
    select(-c(emissions_co2_mt, emissions_ch4_mt, emissions_n2o_mt, emissions_nox_mt, emissions_sox_mt, emissions_co_mt, emissions_vocs_mt, emissions_pm2_5_mt, emissions_pm10_mt))
  
  # 2.2 Distribute broadcasting for countries with catch ASSUMPTION: 1 ton = same emissions for all species
  total_flag_catch <- sau_catch_summary %>%
    filter(zone == regions[i]) %>%
    group_by(zone, year, flag, species) %>%
    summarise(country_total_tons_by_species = sum(sau_catch_tons, na.rm = TRUE), .groups = "drop") %>%
    ungroup() %>%
    group_by(zone, year, flag) %>%
    mutate(country_total_tons = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
    mutate(prop_species_catch = country_total_tons_by_species/country_total_tons) %>%
    filter(country_total_tons_by_species > 0) %>%
    select(-c(country_total_tons)) %>% #country_total_tons_by_species
    filter(!is.na(prop_species_catch))
  
  # TABLE 2
  total_region_broad_allocation <- full_join(total_flag_catch, emissions_partitioned, by = c("zone", "year", "flag")) %>%
    mutate(dist_broad_co2_mt = ifelse(is.na(prop_species_catch), broad_co2_mt, prop_species_catch * broad_co2_mt),
           dist_broad_ch4_mt = ifelse(is.na(prop_species_catch), broad_ch4_mt, prop_species_catch * broad_ch4_mt),
           dist_broad_n2o_mt = ifelse(is.na(prop_species_catch), broad_n2o_mt, prop_species_catch * broad_n2o_mt),
           dist_broad_nox_mt = ifelse(is.na(prop_species_catch), broad_nox_mt, prop_species_catch * broad_nox_mt),
           dist_broad_sox_mt = ifelse(is.na(prop_species_catch), broad_sox_mt, prop_species_catch * broad_sox_mt),
           dist_broad_co_mt = ifelse(is.na(prop_species_catch), broad_co_mt, prop_species_catch * broad_co_mt),
           dist_broad_vocs_mt = ifelse(is.na(prop_species_catch), broad_vocs_mt, prop_species_catch * broad_vocs_mt),
           dist_broad_pm2_5_mt = ifelse(is.na(prop_species_catch), broad_pm2_5_mt, prop_species_catch * broad_pm2_5_mt),
           dist_broad_pm10_mt = ifelse(is.na(prop_species_catch), broad_pm10_mt, prop_species_catch * broad_pm10_mt)) %>%
    select(-country_total_tons_by_species)
  
  
  # -------- TABLE 3 --------
  # TABLE 3: Full join Tables 1 and 2 (to not lose flags with non-broad emissions but no broad emissions)
  total_region_emissions <- full_join(total_region_broad_allocation, total_region_non_broad_allocation, by = c("zone", "year", "flag", "species")) %>%
    #filter(zone == 18) %>%
    # Sum broadcasting and non-broadcasting emissions
    mutate(total_co2_mt = rowSums(across(c(dist_broad_co2_mt, dist_non_broad_co2_mt)), na.rm = TRUE),
           total_ch4_mt = rowSums(across(c(dist_broad_ch4_mt, dist_non_broad_ch4_mt)), na.rm = TRUE),
           total_n2o_mt = rowSums(across(c(dist_broad_n2o_mt, dist_non_broad_n2o_mt)), na.rm = TRUE),
           total_nox_mt = rowSums(across(c(dist_broad_nox_mt, dist_non_broad_nox_mt)), na.rm = TRUE),
           total_sox_mt = rowSums(across(c(dist_broad_sox_mt, dist_non_broad_sox_mt)), na.rm = TRUE),
           total_co_mt = rowSums(across(c(dist_broad_co_mt, dist_non_broad_co_mt)), na.rm = TRUE),
           total_vocs_mt = rowSums(across(c(dist_broad_vocs_mt, dist_non_broad_vocs_mt)), na.rm = TRUE),
           total_pm2_5_mt = rowSums(across(c(dist_broad_pm2_5_mt, dist_non_broad_pm2_5_mt)), na.rm = TRUE),
           total_pm10_mt = rowSums(across(c(dist_broad_pm10_mt, dist_non_broad_pm10_mt)), na.rm = TRUE))
  
  # Keep desired columns
  total_region_emissions <- total_region_emissions %>%
    select(zone,
           year,
           flag,
           species,
           country_total_tons_by_species,
           total_co2_mt,
           total_ch4_mt,
           total_n2o_mt,
           total_nox_mt,
           total_sox_mt,
           total_co_mt,
           total_pm2_5_mt,
           total_pm10_mt
    )
  
  # Account for emissions after full join
  before <- sum(emissions_partitioned$broad_co2_mt, na.rm = TRUE) + sum(sau_summary_non_broadcasting$non_broad_co2_mt, na.rm = TRUE)
  after <- sum(total_region_emissions$total_co2_mt, na.rm = TRUE) 
  
  percent_diff <- ((after-before)/before)*100
  
  # Warning if emissions loss is greater than 0.001%
  if(percent_diff > 0.001) {
    warning(paste0("Greater than 0.001% of emissions lost during full join of broadcasting and non-broadcasting in Region ", regions[i], ". Check Table 3."))
  }
  
  # Append to vector
  option_1[[i]] <- total_region_emissions
}

# Full emissions dataset
full_emissions_sau <- bind_rows(option_1)

# Save SAU emissions dataset to workbench-2
write_csv(full_emissions_sau, file.path("/capstone/seamissions/checkpoint/full_emissions_sau.csv"))
```

# Compare Results

## Load Data

```{r}
#| code-summary: Load Data
#full_emissions_fao <- read_csv(file.path("/capstone/seamissions/checkpoint/full_emissions_fao.csv"), show_col_types = FALSE)

#full_emissions_sau <- read_csv(file.path("/capstone/seamissions/checkpoint/full_emissions_sau.csv"), show_col_types = FALSE)
```

## Join FAO Species Key

[The current International Standard Statistical Classification of Aquatic Animals and Plants (ISSCAAP) in use from 2000](https://www.fao.org/fishery/docs/DOCUMENT/cwp/handbook/annex/AnnexS2listISSCAAP2000.pdf)

```{r}
#| code-summary: Join Species Key
# Import species key
full_species_key <- read_csv(file.path("/capstone/seamissions/data/full_species_key.csv"), show_col_types = FALSE)

# Join to FAO data
full_emissions_fao_species <- left_join(full_emissions_fao, full_species_key, by = c("species_identifier" = "identifier")) %>%
  relocate(c(name_en, scientific_name, major_group, isscaap_group, isscaap), .after = flag)
 
# Save species-specific emissions
write_csv(full_emissions_fao_species, file.path("/capstone/seamissions/checkpoint/full_emissions_fao_species.csv"))
```

## Overall Comparisons

```{r}
#| code-summary: Overall
# -------- Total emissions comparison --------
# Make sure total emissions match
total_co2 <- sum(emissions_partitioned_grouped$co2, na.rm = TRUE)
total_co2_fao <- sum(full_emissions_fao$total_co2_mt, na.rm = TRUE)
total_co2_sau <- sum(full_emissions_sau$total_co2_mt, na.rm = TRUE)

if (total_co2_fao != total_co2_sau) {
  warning("Total emissions in FAO do not match total emissions in SAU.")
}

# -------- Total catch comparison --------
# Filter for years with FAO/SAU overlap (2016-2019)
full_emissions_fao_filtered <- full_emissions_fao %>%
  filter(year <= 2019)

full_emissions_sau_filtered <- full_emissions_sau %>%
  filter(year <= 2019)

# Compare total catch to determine underreporting
fao_total_catch <- sum(full_emissions_fao_filtered$country_total_tons_by_species, na.rm = TRUE)
sau_total_catch <- sum(full_emissions_sau_filtered$country_total_tons_by_species, na.rm = TRUE)

# -------- Compare Emissions/Unit Catch --------
fao_total_emissions_catch <- total_co2_fao/fao_total_catch
sau_total_emissions_catch <- total_co2_sau/sau_total_catch
```

## Emissions/Unit Catch

### Flag

```{r}
#| code-summary: Flag
# Summarise FAO by flag
fao_summary <- full_emissions_fao_filtered %>%
  group_by(flag) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            fao_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(fao_emissions_unit_catch = ifelse(fao_total_catch == 0, 0, fao_total_co2_mt/fao_total_catch)) 
  #select(-c(fao_total_co2_mt, fao_total_catch))

# Summarise SAU by flag
sau_summary <- full_emissions_sau_filtered %>%
  group_by(flag)  %>%
  summarise(sau_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            sau_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(sau_emissions_unit_catch = ifelse(sau_total_catch == 0, 0, sau_total_co2_mt/sau_total_catch)) 
  #select(-c(sau_total_co2_mt, sau_total_catch))

# Join summaries
flag_emissions_catch_comparison <- full_join(fao_summary, sau_summary, by = "flag") 

flag_emissions_catch_comparison %>%
  # Remove flag NA
  filter(!is.na(flag)) %>%
  pivot_longer(cols = 2:3, names_to = "entity", values_to = "emissions_catch") %>%
  arrange(desc(emissions_catch)) %>%
  # Top flags
  filter(flag %in% c("FLK", "URY", "ISR", "PCN", "ITA", "BRA", "SHN", "REU", "CYM", "BEL")) %>%
  ggplot(aes(reorder(flag, -emissions_catch), emissions_catch, fill = entity)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  labs(x = "Flag",
       y = "Emissions/Unit Catch",
       title = "Emissions per Unit Catch by Flag",
       fill = "Agency") +
  theme_minimal()

#write_csv(flag_emissions_catch_comparison, file.path("/capstone/seamissions/checkpoint/flag_emissions_catch_comparison.csv"))
```

### Zone

```{r}
#| code-summary: Region
# Summarise FAO catch by region
fao_summary <- full_emissions_fao_filtered %>%
  group_by(zone) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            fao_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(fao_emissions_unit_catch = ifelse(fao_total_catch == 0, 0, fao_total_co2_mt/fao_total_catch)) %>%
  select(zone, fao_emissions_unit_catch)

sau_summary <- full_emissions_sau_filtered %>%
  group_by(zone) %>%
  summarise(sau_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            sau_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(sau_emissions_unit_catch = ifelse(sau_total_catch == 0, 0, sau_total_co2_mt/sau_total_catch)) %>%
  select(zone, sau_emissions_unit_catch)

# Join summaries
zone_emissions_catch_comparison <- full_join(fao_summary, sau_summary, by = "zone") %>%
  pivot_longer(cols = 2:3, names_to = "entity", values_to = "emissions_catch")

zone_emissions_catch_comparison %>%
  # Order zones by disparity
  ggplot(aes(x = fct_reorder(as.character(zone), -emissions_catch), y = emissions_catch, fill = entity)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  labs(x = "Major Fishing Area",
       y = "Emissions per Unit Catch",
       fill = "Agency",
       title = "Emissions per Unit Catch Disparity by Major Fishing Area") +
  theme_minimal()

zone_emissions_catch_comparison %>%
  arrange(desc(emissions_catch))

write_csv(zone_emissions_catch_comparison, file.path("/capstone/seamissions/checkpoint/zone_emissions_catch_comparison.csv"))

zone_18_fao <- full_emissions_fao_filtered %>%
  filter(zone == 18) %>%
  group_by(flag) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            fao_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(fao_emissions_unit_catch = ifelse(fao_total_catch == 0, 0, fao_total_co2_mt/fao_total_catch))
  #select(flag, fao_emissions_unit_catch)
  
zone_18_sau <- full_emissions_sau_filtered %>%
  filter(zone == 18) %>%
  group_by(flag) %>%
  summarise(sau_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            sau_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(sau_emissions_unit_catch = ifelse(sau_total_catch == 0, 0, sau_total_co2_mt/sau_total_catch))
  #select(flag, sau_emissions_unit_catch)

zone_18 <- full_join(zone_18_fao, zone_18_sau, by = c("flag"))

write_csv(zone_18, file.path("/capstone/seamissions/checkpoint/zone_18.csv"))
```

### Zone and Flag

```{r}
# code-summary: Zone and Flag
# Summarise FAO data
fao_summary <- full_emissions_fao_filtered %>%
  group_by(flag, species_identifier) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            fao_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(fao_emissions_unit_catch = ifelse(fao_total_catch == 0, 0, fao_total_co2_mt/fao_total_catch)) %>%
  select(-c(fao_total_co2_mt, fao_total_catch))

# Summarise SAU data
sau_summary <- full_emissions_sau_filtered %>%
  group_by(flag, species) %>%
  summarise(sau_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            sau_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(sau_emissions_unit_catch = ifelse(sau_total_catch == 0, 0, sau_total_co2_mt/sau_total_catch)) %>%
  select(-c(sau_total_co2_mt, sau_total_catch))

# Join summaries
zone_flag_emissions_catch <- full_join(fao_summary, sau_summary, by = c("flag", "species_identifer"))


zone_flag_emissions_catch %>%
  pivot_longer(cols = 3:4, names_to = "entity", values_to = "emissions_catch") %>%
  filter(flag %in% c("FLK", "URY", "ISR", "PCN", "ITA", "SHN", "REU", "BRA", "CYM", "BEL")) %>%
ggplot(aes(reorder(flag, -emissions_catch), emissions_catch, fill = entity)) +
  geom_boxplot(position = position_dodge2(preserve = "single")) +
  scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  labs(x = "Flag",
       y = "Emissions CO2/Unit Catch",
       fill = "Agency",
       title = "Emissions CO2 per Unit Catch by Flag") +
  theme_minimal()

# VISUALIZE
```

### Flag and Species 

```{r}
# code-summary: Zone and Flag
# Summarise FAO data
fao_summary <- full_emissions_fao_species %>%
  group_by(flag, isscaap_group) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            fao_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(fao_emissions_unit_catch = ifelse(fao_total_catch == 0, 0, fao_total_co2_mt/fao_total_catch)) %>%
  select(-c(fao_total_co2_mt, fao_total_catch))

# Summarise SAU data
sau_summary <- full_emissions_sau_filtered %>%
  group_by(flag, species) %>%
  summarise(sau_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            sau_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(sau_emissions_unit_catch = ifelse(sau_total_catch == 0, 0, sau_total_co2_mt/sau_total_catch)) %>%
  select(-c(sau_total_co2_mt, sau_total_catch))

# Join summaries
# zone_flag_emissions_catch <- full_join(fao_summary, sau_summary, by = c("flag", "species_identifer")) 

fao_summary %>%
  #pivot_longer(cols = 3:4, names_to = "entity", values_to = "emissions_catch") %>%
  #filter(flag %in% c("FLK", "URY", "ISR", "PCN", "ITA", "SHN", "REU", "BRA", "CYM", "BEL")) %>%
  group_by(flag) %>%
  mutate(mean = mean(fao_emissions_unit_catch)) %>%
  arrange(desc(mean)) %>%
  filter(!isscaap_group == "no_fao_species_catch_data") %>%
  filter(flag %in% c("GIB", "SVN", "REU", "MLT", "NRU", "BEL", "JPN", "CYP", "ITA", "BHR", "ISR", "HRV")) %>%
ggplot(aes(reorder(flag, -fao_emissions_unit_catch), fao_emissions_unit_catch)) +
  geom_boxplot() +
  #scale_y_continuous(limits = c(0, 65)) +
  #scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  labs(x = "",
       y = "Emissions CO2/Unit Catch",
       title = "Emissions CO2 per Unit Catch by Flag") +
  theme_minimal()
```

```{r}
# Summarise FAO data
fao_summary <- full_emissions_fao_species %>%
  group_by(zone, flag, isscaap_group) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            fao_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(fao_emissions_unit_catch = ifelse(fao_total_catch == 0, 0, fao_total_co2_mt/fao_total_catch)) %>%
  select(-c(fao_total_co2_mt, fao_total_catch))

# # Summarise SAU data
# sau_summary <- full_emissions_sau_filtered %>%
#   group_by(flag, species) %>%
#   summarise(sau_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
#             sau_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
#   mutate(sau_emissions_unit_catch = ifelse(sau_total_catch == 0, 0, sau_total_co2_mt/sau_total_catch)) %>%
#   select(-c(sau_total_co2_mt, sau_total_catch))

# Join summaries
# zone_flag_emissions_catch <- full_join(fao_summary, sau_summary, by = c("flag", "species_identifer")) 

fao_summary %>%
  #pivot_longer(cols = 3:4, names_to = "entity", values_to = "emissions_catch") %>%
  #filter(flag %in% c("FLK", "URY", "ISR", "PCN", "ITA", "SHN", "REU", "BRA", "CYM", "BEL")) %>%
  filter(!isscaap_group == "no_fao_species_catch_data") %>%
ggplot(aes(as.factor(zone), fao_emissions_unit_catch, group = as.factor(zone))) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0,100)) +
  #scale_y_continuous(limits = c(0, 65)) +
  #scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  labs(x = "FAO Zone",
       y = "Emissions CO2/Unit Catch",
       title = "Emissions CO2 per Unit Catch by Zone") +
  theme_minimal()


```



### ISSCAAP Group

```{r}
#| code-summary: ISSCAAP Group
# Filter FAO species data
full_emissions_fao_species_filtered <- full_emissions_fao_species %>%
  filter(year <= 2019)

# Summarise FAO by isscaap group
fao_summary <- full_emissions_fao_species %>%
  group_by(zone, flag, isscaap_group) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE),
            fao_total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(fao_emissions_unit_catch = ifelse(fao_total_catch == 0, 0, fao_total_co2_mt/fao_total_catch)) %>%
  select(-c(fao_total_co2_mt, fao_total_catch))

# Plot
fao_summary %>%
  #filter(!is.na(isscaap_group)) %>%
  #arrange(desc())
  #slice(1:10) %>%
  #filter(fao_emissions_unit_catch < 100) %>%
  filter(isscaap_group %in% c("Carps, barbels, and other cyprinids", "Salmons, trouts, smelts", "King crabs, squat-lobsters", "Sturgeons, paddlefishes", "Krill, planktonic crustaceans", "Cods, hakes, haddocks", "Scallops, pectens", "Corals", "Sea-squirts and other tunicates", "Mussels", "Sponges", "Clams, cockles, arkshells", "Oysters")) %>%
 ggplot(aes(reorder(isscaap_group, fao_emissions_unit_catch), fao_emissions_unit_catch)) +
  geom_boxplot() +
  #geom_text(aes(label = round(fao_emissions_unit_catch, 1)), hjust = 1.5, color = "white") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 35)) +
  #scale_y_continuous(labels = scales::comma) +
  labs(x = "",
       y = "Emissions CO2/Unit Catch",
       title = "Emissions CO2 per Unit Catch by ISSCAAP Group") +
  theme_minimal()
```

## Agency

### Regional Comparison by Flag

```{r}
#| code-summary: Region and Flag
# Summarise FAO data
fao_summary <- full_emissions_fao_filtered %>%
  group_by(zone, flag) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE))

# Summarise SAU data
sau_summary <- full_emissions_sau_filtered %>%
  group_by(zone, flag) %>%
  summarise(sau_total_co2_mt = sum(total_co2_mt, na.rm = TRUE))

# Join summaries
zone_flag_comparison_table <- full_join(fao_summary, sau_summary, by = c("zone", "flag")) %>%
  mutate(percent_reported_co2 = round((fao_total_co2_mt/sau_total_co2_mt)*100, 2),
         tonnage_discrepancy_co2 = round(fao_total_co2_mt - sau_total_co2_mt, 2)) %>%
  group_by(zone) %>%
  arrange(tonnage_discrepancy_co2, .by_group = TRUE)

# Save table
#write_csv(zone_flag_comparison_table, file.path("/capstone/seamissions/checkpoint/zone_flag_comparison_table.csv"))
```

### Regional Comparison by Flag and Year

```{r}
#| code-summary: Region, Flag, and Year
# Summarise FAO data
fao_summary <- full_emissions_fao_filtered %>%
  group_by(zone, year, flag) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE))

# Summarise SAU data
sau_summary <- full_emissions_sau_filtered %>%
  group_by(zone, year, flag) %>%
  summarise(sau_total_co2_mt = sum(total_co2_mt, na.rm = TRUE))

# Join summaries
zone_year_flag_comparison_table <- full_join(fao_summary, sau_summary, by = c("zone", "year", "flag")) %>%
  mutate(percent_reported_co2 =round((fao_total_co2_mt/sau_total_co2_mt)*100, 2),
         tonnage_discrepancy_co2 = round(fao_total_co2_mt - sau_total_co2_mt, 2))  %>%
  group_by(zone, year) %>%
  arrange(tonnage_discrepancy_co2, .by_group = TRUE)

# Save table
#write_csv(zone_year_flag_comparison_table, file.path("/capstone/seamissions/checkpoint/zone_year_flag_comparison_table.csv"))
```

### Flag

```{r}
#| code-summary: Global Flags
# -------- Emissions --------
# Summarise FAO emissions by flag
fao_summary <- full_emissions_fao_filtered %>%
  group_by(flag) %>%
  summarise(fao_total_co2_mt = sum(total_co2_mt, na.rm = TRUE))

# Summarise SAU emissions by flag
sau_summary <- full_emissions_sau_filtered %>%
  group_by(flag) %>%
  summarise(sau_total_co2_mt = sum(total_co2_mt, na.rm = TRUE))

# Join summaries
flag_emissions_comparison <- full_join(fao_summary, sau_summary, by = "flag") %>%
  mutate(percent_reported_co2 = round((fao_total_co2_mt/sau_total_co2_mt)*100, 2),
         tonnage_discrepancy_co2 = round(fao_total_co2_mt - sau_total_co2_mt, 2)) %>%
  arrange(desc(tonnage_discrepancy_co2), .by_group = TRUE)

# Save table
#write_csv(flag_emissions_comparison, file.path("/capstone/seamissions/checkpoint/flag_emissions_comparison.csv"))

flag_emissions_comparison %>%
  select(flag, fao_total_co2_mt, sau_total_co2_mt) %>%
  pivot_longer(cols = 2:3, names_to = "entity", values_to = "co2_emissions_mt") %>%
  arrange(desc(co2_emissions_mt)) %>%
  slice(1:14) %>%
  ggplot(aes(reorder(flag, -co2_emissions_mt), co2_emissions_mt, fill = entity)) +
           geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  labs(x = "Flag",
       y = "CO2 Emissions (metric tons)",
       fill = "Agency",
       title = "Emissions Disparity by Flag") +
  theme_minimal()

# -------- Catch ---------
# Summarise FAO catch by flag
fao_summary <- full_emissions_fao_filtered %>%
  group_by(flag) %>%
  summarise(fao_catch_tons = sum(country_total_tons_by_species, na.rm = TRUE))

# Summarise SAU catch by flag
sau_summary <- full_emissions_sau_filtered %>%
  group_by(flag) %>%
  summarise(sau_catch_tons = sum(country_total_tons_by_species, na.rm = TRUE))

flag_catch_comparison <- full_join(fao_summary, sau_summary, by = "flag") %>%
  mutate(percent_reported_catch = round((fao_catch_tons/sau_catch_tons)*100, 2),
         tonnage_discrepancy_catch = round(fao_catch_tons - sau_catch_tons, 2)) %>%
  arrange(desc(tonnage_discrepancy_catch), .by_group = TRUE)

# Save table
#write_csv(flag_catch_comparison, file.path("/capstone/seamissions/checkpoint/flag_catch_comparison.csv"))

flag_catch_comparison %>%
  select(flag, fao_catch_tons, sau_catch_tons) %>%
  pivot_longer(cols = 2:3, names_to = "entity", values_to = "catch_tons") %>%
  arrange(desc(catch_tons)) %>%
  slice(1:16) %>%
  ggplot(aes(reorder(flag, -catch_tons), catch_tons, fill = entity)) +
           geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  labs(x = "Flag",
       y = "Catch (tons)",
       fill = "Agency",
       title = "Catch Disparity by Flag") +
  theme_minimal()
```

### Region

```{r}
#| code-summary: Region
# Summarise FAO catch by region
fao_summary <- full_emissions_fao_filtered %>%
  group_by(zone) %>%
  summarise(fao_catch_tons = sum(country_total_tons_by_species, na.rm = TRUE))

# Summarise SAU catch by region
sau_summary <- full_emissions_sau_filtered %>%
  group_by(zone) %>%
  summarise(sau_catch_tons = sum(country_total_tons_by_species, na.rm = TRUE))

zone_catch_comparison <- full_join(fao_summary, sau_summary, by = "zone") %>%
  mutate(diff = sau_catch_tons - fao_catch_tons) %>%
  pivot_longer(cols = 2:3, names_to = "entity", values_to = "catch_tons") %>%
  arrange(desc(diff))

zone_catch_comparison %>%
  # Order zones by disparity
  ggplot(aes(x = fct_reorder(as.character(zone), diff, .desc = TRUE), y = catch_tons, fill = entity)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  labs(x = "Major Fishing Area",
       y = "Catch (tons)",
       fill = "Agency",
       title = "Catch Disparity by Major Fishing Area") +
  theme_minimal()
```

### Year

```{r}
#| code-summary: Year
# -------- Catch ---------
# Summarise FAO catch by region
fao_summary <- full_emissions_fao_filtered %>%
  group_by(year) %>%
  summarise(fao_catch_tons = sum(country_total_tons_by_species, na.rm = TRUE))

# Summarise SAU catch by region
sau_summary <- full_emissions_sau_filtered %>%
  group_by(year) %>%
  summarise(sau_catch_tons = sum(country_total_tons_by_species, na.rm = TRUE))

year_catch_comparison <- full_join(fao_summary, sau_summary, by = "year") %>%
  pivot_longer(cols = 2:3, names_to = "entity", values_to = "catch_tons") %>%
  arrange(desc(catch_tons)) %>%
  slice(1:20)

year_catch_comparison%>%
  ggplot(aes(as.factor(year), catch_tons, fill = entity)) +
           geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(labels = c("FAO", "SAU"), values = c("cornflowerblue", "firebrick")) +
  labs(x = "Zone",
       y = "Catch (tons)",
       fill = "Agency",
       title = "Catch Disparity by Year") +
  theme_minimal()
```

## Visualizations

```{r}
#| code-summary: Visualize

# Graph by ISSCAAP group
full_emissions_fao_species %>%
  # Remove emissions from "no-fao-species-catch-data"
  filter(!isscaap == 99999999) %>%
  # Group by ISSCAAP group
  group_by(isscaap_group) %>%
  summarise(co2 = sum(total_co2_mt, na.rm = TRUE)) %>%
  filter(!is.na(isscaap_group)) %>%
ggplot(aes(reorder(as.factor(isscaap_group), co2), co2)) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  coord_flip()+
  labs(x = "ISSCAAP Groups",
    y = "CO2 Emissions (mt)") +
  theme_minimal()

# Graph by ISSCAAP group
full_emissions_fao_species %>%
  # Remove emissions from "no-fao-species-catch-data"
  filter(!isscaap == 99999999) %>%
  # Group by ISSCAAP group
  group_by(isscaap_group) %>%
  summarise(total_emissions = sum(total_co2_mt, na.rm = TRUE),
            total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(emissions_unit_catch = total_emissions/total_catch) %>%
ggplot(aes(reorder(as.factor(isscaap_group), emissions_unit_catch), emissions_unit_catch)) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  coord_flip()+
  labs(x = "ISSCAAP Groups",
    y = "CO2 Emissions/Unit Catch") +
  theme_minimal()

# Graph by ISSCAAP group
full_emissions_fao_species %>%
  # Remove emissions from "no-fao-species-catch-data"
  filter(!isscaap == 99999999) %>%
  # Group by ISSCAAP group
  group_by(year, isscaap_group) %>%
  summarise(total_emissions = sum(total_co2_mt, na.rm = TRUE),
            total_catch = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  mutate(emissions_unit_catch = total_emissions/total_catch) %>%
ggplot(aes(year, emissions_unit_catch, color = isscaap_group)) +
  geom_line() +
  gghighlight(mean(emissions_unit_catch) > 2.6) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "",
    y = "CO2 Emissions/Unit Catch") +
  theme_minimal() +
  theme(plot.margin = margin(1, 5, 1, 1, "cm"), 
        plot.title.position = "plot",
    plot.caption.position = "plot",
        legend.position = "bottom",
    panel.spacing = unit(1, "lines"),
        legend.title = element_blank(),
        clip = "off")
  
theme(
    plot.margin = margin(1, 5, 1, 1, "cm"),  # extra right margin
    plot.title.position = "plot",
    plot.caption.position = "plot",
    panel.spacing = unit(1, "lines"),
    clip = "off"  # <--- IMPORTANT
  )

# Emissions per unit catch over time for a group (facet region?)
full_emissions_fao_species %>%
  # Remove emissions from "no-fao-species-catch-data"
  filter(!isscaap == 99999999) %>%
  group_by(year, isscaap_group) %>%
  summarise(emissions_unit_catch = sum(total_co2, na.rm = TRUE)/sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  arrange(desc(emissions_unit_catch)) %>%
  slice(1:6) %>%
  ggplot(aes(year, emissions_unit_catch, color = isscaap_group)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "",
    y = "CO2 Emissions (tons)/Unit Catch (tons)",
    color = "ISSCAAP Group",
    title = "Top 5 ISSCAAP Groups by Emissions/Unit Catch") +
  theme_minimal()
  
# Emissions/unit catch ISSCAAP Group by Region over time
full_emissions_fao_species %>%
  # Remove emissions from "no-fao-species-catch-data"
  filter(!isscaap == 99999999) %>%
  group_by(zone, year, isscaap_group) %>%
  summarise(emissions_unit_catch = sum(total_co2, na.rm = TRUE)/sum(country_total_tons_by_species, na.rm = TRUE)) %>%
  #arrange(desc(emissions_unit_catch)) %>%
  slice_max(order_by = desc(emissions_unit_catch), n = 3) %>%
  ggplot(aes(year, emissions_unit_catch, color = isscaap_group)) +
  geom_line() +
  facet_wrap(~zone, scales = "free_y") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "",
    y = "CO2 Emissions (mt)",
    color = "",
    title = "Top 3 ISSCAAP Groups by Emissions/Unit Catch per Major Fishing Area") +
  theme_minimal() +
  theme(legend.position = "bottom")


# For a specific zone:

# For a specific isscaap group:



```
