---
title: "Visualizations"
subtitle: "Flash Talk 3"
author: "Carmen Hoyt"
format:
  html:
      toc: true
      code-fold: true
editor: visual
execute:
  warning: false
  messages: false
editor_options: 
  chunk_output_type: console
---

# Visualizations

Visualizing the emissions dataset for Flash Talk 3.

```{r}
#| code-summary: Expand code
library(tidyverse)
library(janitor)
library(here)
library(gghighlight)
library(lubridate)
library(ggwordcloud)
library(sf)
library(raster)
library(tmap)
library(treemapify)
library(paletteer)
```

## Load Data

```{r}
#| code-summary: Expand code

# Read in emissions data
emissions <- read_csv(here("checkpoint", "emissions.csv"), show_col_types = FALSE)
em_pivot <- read_csv(here("checkpoint", "em_pivot.csv"), show_col_types = FALSE)

# Read in FAO shapefile
#shp <- st_read('/capstone/seamissions/data/fao_region_shapefile/World_Fao_Zones.shp')
```

### Word Clouds

```{r}
#| code-summary: Expand code

# Word cloud of pollutants by total emissions (all time)
em_pivot %>%
  group_by(pollutant_type) %>%
  summarise(emissions = sum(emissions)) %>%
ggplot(aes(label = pollutant_type, size = emissions)) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 15) +  # Increase max_size for bigger words
  theme_minimal()

# Word cloud of flags (all time emissions)
em_pivot %>%
  group_by(flag) %>%
  summarise(emissions = sum(emissions)) %>%
ggplot(aes(label = flag, size = emissions)) +
  geom_text_wordcloud() +
  theme_void() +
  theme(
  text = element_text(family = "Comic Sans MS")
  )
```

### Treemaps

```{r}
#| code-summary: Expand code

# Flag by all time emissions
em_pivot %>%
  mutate(flag = ifelse(flag %in% c("CHN", "RUS", "NOR", "JPN", "KOR", "TWN", "USA"), flag, "other")) %>%
  group_by(flag) %>%
  summarise(emissions = sum(emissions)) %>%
  arrange(desc(emissions)) %>%
  slice(1:10) %>%
  mutate(labels = case_when(
    flag == "CHN" ~ "China",
    flag == "RUS" ~ "Russia", 
    flag == "NOR" ~ "Norway", 
    flag == "JPN" ~ "Japan",
    flag == "KOR" ~ "South Korea", 
    flag == NA ~ "No Flag",
    flag == "TWN" ~ "Taiwan",
    flag == "USA" ~ "United States",
    flag == "ISL" ~ "Israel",
    flag == "ESP" ~ "Spain"
  )) %>%
ggplot(aes(area = emissions, fill = flag, label = labels)) +
  geom_treemap() + 
  geom_treemap_text(color = "white", 
                    place = "center", 
                    #fontface = "bold",
                    grow = TRUE,
                    size = 10) +
  #scale_fill_manual(palette = "Blues") +
  theme_minimal() +
  theme(
    # remove legend
    legend.position = 'none'
    )

# Pollutant type by all time emissions (data)
plot <- em_pivot %>%
  group_by(pollutant_type) %>%
  summarise(emissions = sum(emissions)) %>%
  mutate(labels = case_when(
    pollutant_type == "ch4" ~ "CH4",
    pollutant_type == "co" ~ "CO", 
    pollutant_type == "co2" ~ "CO2", 
    pollutant_type == "n2o" ~ "N2O",
    pollutant_type == "nox" ~ "NOX", 
    pollutant_type == "pm" ~ "PM",
    pollutant_type == "pm10" ~ "PM10",
    pollutant_type == "pm2_5" ~ "PM2.5",
    pollutant_type == "sox" ~ "SOX",
    pollutant_type == "vocs" ~ "VOCs"
  )) 

custom_labels <- setNames(plot$labels, plot$pollutant_type)

# Pollutant type by all time emissions (plot)
ggplot(plot, aes(area = emissions, fill = pollutant_type, label = labels)) +
  geom_treemap() +
  geom_treemap_text(color = "white",
                    place = "center",
                    fontface = "bold",
                    grow = TRUE,
                    size = 10) +
  #scale_fill_brewer(palette = "Dark2") +
  scale_fill_paletteer_d("ggthemes::Classic_Color_Blind", labels = custom_labels) +
  theme_minimal() +
  labs(fill = "Pollutant") +
    theme(legend.position = 'none')

# vessel classes
em_pivot %>%
  mutate(vessels = ifelse(vessel_class %in% c("drifting_longlines", "passenger", "set_longlines", "squid_jigger", "trawlers", "tuna_purse_seines", "pole_and_line", "pots_and_traps", "tug", "set_gillnets"), vessel_class, "other")) %>%
  group_by(vessels) %>%
  summarise(emissions = sum(emissions)) %>%
  mutate(labels = case_when(
    vessels == "drifting_longlines" ~ "Drifting Longlines",
    vessels == "passenger" ~ "Passenger", 
    vessels == "set_longlines" ~ "Set Longlines", 
    vessels == "squid_jigger" ~ "Squid Jigger",
    vessels == "trawlers" ~ "Trawlers", 
    vessels == "tuna_purse_seines" ~ "Tuna Purse Seines",
    vessels == "pole_and_line" ~ "Pole and Line",
    vessels == "pots_and_traps" ~ "Pots and Traps",
    vessels == "tug" ~ "Tug",
    vessels == "set_gillnets" ~ "Set Gillnets",
    vessels == "other" ~ "Other"
  )) %>%
  arrange(desc(emissions)) %>%
  #slice(1:10) %>%
  ggplot(aes(area = emissions, fill = vessels, label = labels)) +
  geom_treemap() +
  geom_treemap_text(color = "black",
                    place = "center",
                    #fontface = "bold",
                    grow = TRUE,
                    size = 10)+
                    #label = labels) +
  scale_fill_brewer(palette = "Set3") +
  # scale_fill_manual(values = c(drifting_longlines = "#FFA508",
  #           passenger = "#034E95",
  #           set_longlines = "#63A545",
  #           squid_jigger = "#008ECC",
  #           trawlers = "#047C90",
  #           tuna_purse_seines = "gray60",
  #           pole_and_line = "brown",
  #           tug = "#e22e12",
  #           set_gillnets = "#00003c",
  #           pots_and_traps = "orange")) +
  #facet_wrap(~length_size_class_percentile) +
  theme_minimal() +
  theme(
    legend.position = 'none'
  )
```

### By vessel class

```{r}
#| code-summary: Expand code

labels <- c(drifting_longlines = "Drifting Longlines",
            passenger = "Passenger", 
            set_longlines = "Set Longlines", 
            squid_jigger = "Squid Jigger",
            trawlers = "Trawlers", 
            tuna_purse_seines = "Tuna Purse Seines",
            pole_and_line = "Pole and Line",
            pots_and_traps = "Pots and Traps",
            tug = "Tug",
            set_gillnets = "Set Gillnets",
            other = "Other"
            )

# emissions of 6 types of vessel classes over time
em_pivot %>%
  filter(pollutant_type == "co2") %>%
  group_by(year, vessel_class, pollutant_type) %>%
  summarise(emissions = sum(emissions)) %>%
  filter(vessel_class %in% c("drifting_longlines", "set_longlines", "passenger", "squid_jigger", "trawlers", "tuna_purse_seines")) %>%
ggplot(aes(year, emissions, color = pollutant_type)) +
  geom_line() +
  facet_wrap(~vessel_class, 
             #scales = "free_y", 
             labeller = labeller(vessel_class = labels)) +
  labs(y = "Emissions (MT)") +
  scale_x_continuous(breaks = seq(min(em_pivot$year), max(em_pivot$year), by = 3)) +
  scale_y_continuous(labels = scales::label_comma(scale = 1, accuracy = 1)) +
  theme_minimal() +
  theme(
    # remove x axis title
    axis.title.x = element_blank(),
    
    # adjust individual plot titles
    strip.text = element_text(size = 12, face = "bold"),
    
    # remove legend
    legend.position = 'none',
    
    # adjust the font and size of the axis text
    axis.text = element_text(family = "sans",
                              size = 10)
    
  )

# emissions by pollutant and vessel class
em_pivot %>%
  group_by(year, vessel_class, pollutant_type) %>%
  summarise(emissions = sum(emissions)) %>%
  filter(vessel_class %in% c("drifting_longlines", "set_longlines", "passenger", "squid_jigger", "trawlers", "tuna_purse_seines")) %>%
ggplot(aes(year, emissions, color = vessel_class)) +
  geom_line() +
  #gghighlight(vessel_class == "trawlers", use_group_by = FALSE) +
  facet_wrap(~pollutant_type, scales = "free_y")


em_pivot %>%
  mutate(vessels = ifelse(vessel_class %in% c("drifting_longlines", "passenger", "set_longlines", "squid_jigger", "trawlers", "tuna_purse_seines", "pole_and_line", "pots_and_traps", "tug", "set_gillnets"), vessel_class, "other")) %>%
  group_by(vessels) %>%
  summarise(emissions = sum(emissions)) %>%
  mutate(labels = case_when(
    vessels == "drifting_longlines" ~ "Drifting Longlines",
    vessels == "passenger" ~ "Passenger", 
    vessels == "set_longlines" ~ "Set Longlines", 
    vessels == "squid_jigger" ~ "Squid Jigger",
    vessels == "trawlers" ~ "Trawlers", 
    vessels == "tuna_purse_seines" ~ "Tuna Purse Seines",
    vessels == "pole_and_line" ~ "Pole and Line",
    vessels == "pots_and_traps" ~ "Pots and Traps",
    vessels == "tug" ~ "Tug",
    vessels == "set_gillnets" ~ "Set Gillnets",
    vessels == "other" ~ "Other"
  )) %>%
  arrange(desc(emissions)) %>%
  ggplot(aes(reorder(vessels, emissions), emissions)) +
  geom_col() +
  coord_flip()
```

### By Flag

```{r}
#| code-summary: Expand code

flag_labels <- c(CHN = "China",
                 JPN = "Japan",
                 KOR = "South Korea",
                 NOR = "Norway",
                 RUS = "Russia",
                 USA = "United States")

# top 5 + usa with 6 vessel types
em_pivot %>%
  group_by(year, flag, vessel_class) %>%
  summarise(emissions = sum(emissions)) %>%
  filter(flag %in% c("CHN", "RUS", "NOR", "JPN", "KOR", "USA")) %>%
  filter(vessel_class %in% c("trawlers", "passenger", "tuna_purse_seines", "drifting_longlines", "set_longlines", "squid_jigger", "pole_and_line")) %>%
  ggplot(aes(year, emissions, color = vessel_class)) +
  geom_line() +
  scale_color_manual(values = c("#1b4b87", "#8abbc7", "#e22e12", "#63A545", "#fcb900", "#7bdcb5", "#32373c"), labels = labels) +
  #scale_color_brewer(palette = "Set3") +
  #gghighlight(vessel_class == "trawlers", use_group_by = FALSE) +
  facet_wrap(~flag,
             labeller = labeller(flag = flag_labels),
             scales = "free_y") +
  labs(y = "Emissions (MT)",
       color = "Vessel Class") +
  scale_x_continuous(breaks = seq(min(em_pivot$year), max(em_pivot$year), by = 3)) +
  scale_y_continuous(labels = scales::label_comma(scale = 1, accuracy = 1)) +
  theme_minimal() +
  theme(
    # remove x axis title
    axis.title.x = element_blank(),
    
    # adjust individual plot titles
    strip.text = element_text(size = 12, face = "bold"),
    
    # remove legend
    #legend.position = 'none',
    
    # adjust the font and size of the axis text
    axis.text = element_text(family = "sans",
                              size = 10)
  )
```

### By spatial extent

```{r}
#| code-summary: Expand code
# Mean global co2
em_pivot %>%
  filter(pollutant_type == "co2") %>%
  group_by(lat_bin, lon_bin) %>%
  summarise(emissions = mean(emissions)) %>%
  mutate(opacity = scales::rescale(emissions, to = c(0, 1))) %>%
  ggplot(aes(x = lon_bin, y = lat_bin, fill = opacity)) +
  geom_raster() +  # Faster rendering for large data
  scale_fill_viridis_c() +
  guides(fill = guide_colorbar(barwidth = 15, 
                               barheight = 0.75,
                               title.position = "top")) +
  theme_minimal() +
  labs(title = "Average Global CO2 Emissions",
       subtitle = "2015-2024",
       fill = "Relative Emissions") +
  theme(
    axis.title = element_blank(),
    
    plot.title = element_text(family = "sans",
                              size = 12,
                              hjust = 0.5),
    
    plot.subtitle = element_text(family = "sans",
                                size = 10,
                                hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_text(family = "sans",
                                size = 10,
                                hjust = 0.5)
  ) 

```

```{r}
test <- broadcasting %>%
  pivot_longer(10:19,
                names_to = "pollutant_type",
                values_to = "emissions") %>%
  mutate(pollutant_type = str_remove(pollutant_type, "emissions_")) %>%
  group_by(year_month, lat_bin, lon_bin) %>%
  summarise(emissions_broad = sum(emissions))

test2 <- non_broadcasting %>%
  pivot_longer(5:14,
                names_to = "pollutant_type",
                values_to = "emissions") %>%
  mutate(pollutant_type = str_remove(pollutant_type, "emissions_")) %>%
  group_by(year_month, lat_bin, lon_bin) %>%
  summarise(emissions_non = sum(emissions))

test_joined <- left_join(test, test2, by = c("year_month", "lat_bin", "lon_bin"))

viz <- test_joined %>%
  group_by(lat_bin, lon_bin) %>%
  summarise(emissions_broad = sum(emissions_broad, na.rm = TRUE),
            emissions_non = sum(emissions_non, na.rm = TRUE)) %>%
  mutate(prop = emissions_non/emissions_broad)

# if (nrow(test) != nrow(test2)) {
#   warning("Mismatched number of pixels. Cannot calculate proportion of ais to non-broadcasting.")
# }

viz %>%
  filter(prop >= 20)

ggplot(viz, aes(x = lon_bin, y = lat_bin, fill = prop)) +
  geom_raster() +  # Faster rendering for large data
  scale_fill_viridis_c(labels = scales::label_percent(scale = 100, big.mark = ",")) +
  guides(fill = guide_colorbar(barwidth = 25, 
                               barheight = 0.75,
                               title.position = "top")) +
  theme_void(base_size = 18) +
  labs(title = "Proportion of Non-Broadcasting to\nBroadcasting Emissions",
       subtitle = "2015-2024",
       fill = "") +
  theme(
    axis.title = element_blank(),
    
    plot.title = element_text(family = "sans",
                              size = rel(1.2),
                              hjust = 0.5),
    
    plot.subtitle = element_text(family = "sans",
                                size = rel(1.0),
                                hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_blank()
    
    # text(family = "sans",
    #                             size = rel(1.0),
    #                             hjust = 0.5)
  ) 

test2 %>%
  group_by(lat_bin, lon_bin) %>%
  summarise(emissions_non = sum(emissions_non)) %>%
ggplot(aes(x = lon_bin, y = lat_bin, fill = emissions_non)) +
  geom_raster() +  # Faster rendering for large data
  scale_fill_viridis_c() +
  guides(fill = guide_colorbar(barwidth = 15, 
                               barheight = 0.75,
                               title.position = "top")) +
  theme_minimal() +
  labs(title = "Non-Broadcasting Emissions",
       subtitle = "",
       fill = "Proportion") +
  theme(
    axis.title = element_blank(),
    
    plot.title = element_text(family = "sans",
                              size = 12,
                              hjust = 0.5),
    
    plot.subtitle = element_text(family = "sans",
                                size = 10,
                                hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_text(family = "sans",
                                size = 10,
                                hjust = 0.5)
  ) 
```


df_sf <- df_sf %>%
  filter(pollutant_type == "co2") %>%
  group_by(geometry) %>%
  summarise(emissions = mean(emissions)) %>%
  mutate(opacity = scales::rescale(emissions, to = c(0, 1))) %>%
  ggplot(aes(fill = opacity)) +
  geom_sf() +
  scale_fill_viridis_c() +
  geom_sf(data = shp, color = "black", fill = NA, size = 0.5) +  # Shapefile overlay
  coord_sf() +
  guides(fill = guide_colorbar(barwidth = 15, barheight = 0.75,
                               title.position = "top")) +
  theme_minimal() +
  labs(title = "Average Global CO2 Emissions",
       subtitle = "2015-2024",
       fill = "Relative Emissions") +
  theme(
    axis.title = element_blank(),

    plot.title = element_text(family = "sans",
                              size = 12,
                              hjust = 0.5),

    plot.subtitle = element_text(family = "sans",
                                size = 10,
                                hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_text(family = "sans",
                                size = 10,
                                hjust = 0.5)
  )
# 
df_sf <- st_as_sf(em_pivot, coords = c("lon_bin", "lat_bin"), crs = 4326)
shp <- st_transform(shp, crs(df_sf))
colnames(shp)

ggplot() +
  # Assuming shapefile has a column "population" for color mapping

  geom_sf(data = df_sf, aes(color = opacity)) +
   geom_sf(data = shp, color = "black", size = 0.5) +
  scale_fill_viridis_c() +  # Apply color scale (viridis for continuous data)
  theme_minimal()

shp <- st_make_valid(shp)
st_crs(shp) == st_crs(df_sf)

invalid_geometries <- st_is_valid(shp)
shp <- shp[invalid_geometries, ]

tm_shape(df_sf) +
  tm_dots(col = "opacity", palette = "viridis") +  # Raster layer with color palette
tm_shape(shp) +
  tm_borders() 

```{r}
#| code-summary: Expand code
em_pivot %>%
  ggplot(aes(as.factor(length_size_class_percentile), emissions)) +
  geom_boxplot()
```

```{r}
#| code-summary: Expand code
em_pivot %>%
  filter(pollutant_type == "co2") %>%
  group_by(length_size_class_percentile, pollutant_type) %>%
  summarise(emissions = mean(emissions)) %>%
  ggplot(aes(pollutant_type, emissions, fill = pollutant_type)) +
  geom_col() +   # Create bars
  coord_polar(theta = "y") + 
  facet_wrap(~length_size_class_percentile) +
  labs(fill = "Pollutant",
       title = "Mean Emissions (Mt) by Vessel Length Class Size Percentile") +
  theme_minimal()

em_pivot %>%
  group_by(length_size_class_percentile, vessel_class) %>%
  summarise(vessels = n()) %>%
  ggplot(aes(vessel_class, vessels, fill = vessel_class)) +
  geom_col() +
  coord_polar(theta = "y") +
  facet_wrap(~length_size_class_percentile) +
  theme_minimal()

em_pivot %>%
  group_by(length_size_class_percentile,  vessel_class) %>%
  summarise(vessels = n()) %>%
  ggplot(aes(length_size_class_percentile, fill = vessel_class)) +
  geom_bar(stat = "count")

em_pivot %>%
  group_by(length_size_class_percentile,  vessel_class) %>%
  summarise(vessels = n()) %>%
  group_by(length_size_class_percentile) %>%
  mutate(total_vessels = sum(vessels)) %>%
  mutate(prop = vessels/total_vessels) %>%
  
  ggplot(aes(x = as.factor(length_size_class_percentile), y = prop, fill = vessel_class)) +
  geom_col(stat = "identity")

# df <- em_pivot %>%
#   group_by(length_size_class_percentile, pollutant_type) %>%
#   summarise(emissions = mean(emissions))
# 
# # Compute fractions of the whole
# df$fraction <- df$emissions / sum(df$emissions)
# df$ymax <- cumsum(df$fraction)
# df$ymin <- c(0, head(df$ymax, n = -1))
# 
# ggplot(df, aes(ymax = ymax, ymin = ymin, xmax = 2, xmin = 1, fill = pollutant_type)) +
#   geom_rect() +  # Create the donut segments
#   geom_text(aes(x = 1.5, y = (ymin + ymax) / 2, label = paste0(round(fraction * 100), "%")), size = 5) +  # Labels
#   coord_polar(theta = "y", start = -pi/2) + # Convert to semi-circle
#   xlim(c(0, 2)) +  # Limit to half-circle
#   facet_wrap(~length_size_class_percentile) +
#   theme_void() +  # Remove grid and axis
#   labs(title = "Fractions of a Whole in a Semi-Circle Donut Chart") +
#   theme(legend.position = "bottom")

```

emLab colors:

```{r}
"#FFA508"

"#034E95"

"#63A545"

"#008ECC"

"#047C90"

"#EDF0F2"

```

GFW colors:

```{r}
"#1b4b87"

"#8abbc7"

"#e22e12"

"#00003c"

"#f78da7"

"#cf2e2e"

"#ff6900"

"#fcb900"

"#7bdcb5"

"#00d084"

"#8ed1fc"

"#0693e3"

"#9b51e0"

"#32373c"
```

Old work: ?!?

## Visualize (maps, time series plots)

#### Vessel focus

```{r}
vessels <- emissions %>%
  mutate(year = as.integer(substring(year_month, 1, 4))) %>%
  group_by(year, vessel_class) %>%
  summarise(co2 = sum(total_co2)) 

ggplot(vessels, aes(year, co2, color = vessel_class)) +
  geom_line() +
  #scale_x_date(labels = scales::date_format("%Y")) +
  scale_x_continuous(breaks = seq(min(vessels$year), max(vessels$year), by = 1)) +
  gghighlight(vessel_class == "trawlers", use_group_by = FALSE)

vessels2 <- emissions %>%
  mutate(year = as.numeric(substring(year_month, 1, 4))) %>%
  group_by(year, vessel_class) %>%
  summarise(co2 = sum(total_co2),
            ch4 = sum(total_ch4),
            n2o = sum(total_n2o),
            nox = sum(total_nox),
            sox = sum(total_sox),
            pm = sum(total_pm),
            co = sum(total_co)) %>%
  pivot_longer(cols = c(3:9), 
               names_to = "pollutant_type",
               values_to = "emissions")

# vessels2$pollutant_type <- factor(vessels2$pollutant_type,
#                           levels = c("co2", "ch4", "n2o"), 
#                           labels = c("CO2", "CH4", "N2O"))

ggplot(vessels2, aes(year, emissions, color = pollutant_type)) +
  geom_line() +
  facet_wrap(~vessel_class, scales = "free_y")
  #gghighlight(pollutant_type == "co2", use_group_by = FALSE)

ggplot(vessels2, aes(year, emissions, color = vessel_class)) +
  geom_line() +
  facet_wrap(~pollutant_type, scales = "free_y")

vessels2 %>%
  filter(pollutant_type == "co2") %>%
  #filter(vessel_class %in% c("drifting_longlines", "passenger", "trawlers", "tuna_purse_seines", "squid_jigger")) %>%
ggplot(aes(year, emissions)) +
  geom_line() +
  facet_wrap(~vessel_class 
             #scales = "free_y"
             ) +
  scale_x_continuous(breaks = seq(min(vessels$year), max(vessels$year), by = 2)) 

vessels2 %>%
  filter(pollutant_type == "co2") %>%
  filter(vessel_class %in% c("drifting_longlines", "passenger", "trawlers", "tuna_purse_seines", "squid_jigger", "pole_and_line")) %>%
ggplot(aes(year, emissions, color = pollutant_type)) +
  geom_line() +
  facet_wrap(~vessel_class, scales = "free_y") +
  scale_x_continuous(breaks = seq(min(vessels$year), max(vessels$year), by = 2))
  #gghighlight(pollutant_type == "co2", use_group_by = TRUE)

vessels2 %>%
  filter(pollutant_type != "co2") %>%
  filter(vessel_class %in% c("drifting_longlines", "passenger", "trawlers", "tuna_purse_seines", "squid_jigger", "pole_and_line")) %>%
ggplot(aes(vessel_class, emissions, fill = pollutant_type)) +
  geom_col(position = position_dodge()) +
  coord_flip()
```

```{r}
flags <- emissions %>%
  mutate(year = as.integer(substring(year_month, 1, 4))) %>%
  group_by(year, flag, vessel_class) %>%
  summarise(co2 = sum(total_co2),
            ch4 = sum(total_ch4),
            n2o = sum(total_n2o),
            nox = sum(total_nox),
            sox = sum(total_sox),
            pm = sum(total_pm),
            co = sum(total_co)) %>%
  pivot_longer(cols = c(4:10), 
               names_to = "pollutant_type",
               values_to = "emissions") 

flags %>%
  filter(pollutant_type == "co2") %>%
  filter(flag %in% c("CHN", "RUS", "NOR", "JPN", "KOR", "USA")) %>%
  filter(vessel_class %in% c("trawlers", "passenger", "patrol_vessel", "tuna_purse_seines", "drifting_longlines", "set_gillnets", "set_longlines")) %>%
ggplot(aes(year, emissions, color = vessel_class)) +
  geom_line() +
  facet_wrap(~flag, scales = "free_y") +
  scale_x_continuous(breaks = seq(min(vessels$year), max(vessels$year), by = 2))
  

#gghighlight(vessel_class == "trawlers", use_group_by = FALSE)


emissions_2017 <- emissions %>%
  mutate(date = lubridate::ym(year_month)) %>%
  mutate(year = year(date)) %>%
  filter(year == "2017")

ggplot(emissions_2017, aes(flag, total_co2)) +
  geom_boxplot() +
  coord_flip()

# em_pivot <- emissions %>%
#   pivot_longer(cols = c(10:16), names_to = "pollutant_type",
#                values_to = "total_emissions_mt") %>%
#   mutate(pollutant_type = str_remove(pollutant_type, "total_"))
# 
# 
# ggplot(em_pivot, aes(pollutant_type, total_emissions_mt)) +
#   geom_boxplot() +
#   coord_flip()

# word cloud of pollutants by total emissions in the dataset (all time)
em_pivot %>%
  group_by(pollutant_type) %>%
  summarise(emissions = sum(total_emissions_mt)) %>%
ggplot(aes(label = pollutant_type, size = emissions)) +
  geom_text_wordcloud() +
  theme_minimal()
```
