---
title: "FAO Regions"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# FAO Regions

### Load packages

```{r}
library(tidyverse)
library(janitor)
library(here)
library(gghighlight)
library(lubridate)
library(ggwordcloud)
library(glue)
library(sf)
library(dplyr)
```

### Import and clean data

```{r}
# Turn off scientific notation
options(scipen=999)

# -------- EMISSIONS --------
#filepath <- paste0(project_directory, "/checkpoint/emissions.csv")

# import emissions data
# workbench-2
emissions <- read_csv(file.path("/capstone/seamissions/checkpoint/emissions.csv"), show_col_types = FALSE)

# local
#emissions <- read_csv(file.path("/Users/carmenhoyt/Documents/UCSB/Capstone/Data/checkpoint/emissions.csv"), show_col_types = FALSE)

# equal earth 8857
#my_crs <- st_crs("+proj=eqearth +datum=WGS84 +units=m +no_defs")

# OR mollweide projection 54009 but not recognized
#mollweide_proj <- "+proj=moll +lon_0=0 +datum=WGS84 +units=m +no_defs"
#emissions_sf <- st_set_crs(emissions_sf, mollweide_proj) # transform?

# filter for desired variables
emissions_filtered <- emissions %>%
  # make year column
  mutate(year = as.integer(substr(year_month, 1, 4))) %>%
  # group by pixel, year, and vessel_class(?)
  group_by(lat_bin, lon_bin, year, flag, vessel_class) %>%
  # sum co2 (ADD OTHER POLLUTANTS HERE)
  summarise(total_co2_mt = sum(total_co2, na.rm = TRUE))

# convert lat/long to point geometry
emissions_sf <- emissions_filtered %>%
  # add lat and lon bin to preserve for merge 
  dplyr::mutate(lat = lat_bin, lon = lon_bin) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) # import as WGS (unit = degrees)

# create grid throughout extent of emissions_sf from point geometry
emissions_grid <- emissions_sf %>%
  st_make_grid(cellsize = c(1,1),
               what = "polygons") %>%
  st_sf() %>%
  st_make_valid()
  
# spatially join emissions data to the grid
emissions_grid_sf <- st_join(emissions_grid, emissions_sf, left = FALSE) %>%
  # transform to Equal Earth
  st_transform(6933)

# add area column before intersection
emissions_grid_sf$area_before <- st_area(emissions_grid_sf)

# -------- FAO REGIONS --------
# import regions data (set same crs)
# workbench-2
fao_regions <- st_read(file.path("/capstone/seamissions/fao_region_shapefile")) %>%
  # transform to same crs as grid
  st_transform(st_crs(emissions_grid_sf)) %>%
  # fix geometries
  st_make_valid()

# local
# fao_regions <- st_read(file.path("/Users/carmenhoyt/Documents/UCSB/Capstone/Data/fao_region_shapefile")) %>%
#   st_transform(st_crs(emissions_grid_sf)) %>%
#   st_make_valid()

# confirm crs match
st_crs(emissions_grid_sf) == st_crs(fao_regions)

# -------- FAO CATCH --------
# import catch data
# workbench-2
fao_catch <- read_csv(file.path("/capstone/seamissions/checkpoint/fao_catch.csv"), show_col_types = FALSE) %>%
  # select desired columns
  select(area_code,
         period,
         iso3_code,
         official_name_en,
         scientific_name,
         value) %>%
  # rename for consistency
  rename(zone = area_code,
         year = period,
         flag = iso3_code,
         catch = value,
         official_name = official_name_en) %>%
  # filter for >2015 to match emissions data
  filter(year >= 2015) %>%
  group_by(year, zone, flag, scientific_name, official_name) %>%
  summarise(total_catch_t = sum(catch, na.rm = TRUE)) %>%
  mutate(zone = as.integer(zone),
         year = as.integer(year))

# local
#fao_catch <- read_csv(file.path("/Users/carmenhoyt/Documents/UCSB/Capstone/Data/checkpoint/fao_catch.csv"), show_col_types = FALSE)

# -------- CHECKPOINT --------
# save locally
#st_write(emissions_grid_sf, here("checkpoint", "emissions_grid_sf.gpkg"))

# save to server
#st_write(emissions_grid_sf, file.path("/capstone/seamissions/checkpoint/emissions_grid_sf.gpkg"), delete_dsn = TRUE)

#st_write(fao_regions, file.path("/capstone/seamissions/checkpoint/fao_regions/fao_regions.shp"), delete_dsn = TRUE)
```

```{r}
# cut out grid cells that overlap with zones (st_intersection)
emissions_zones <- st_intersection(emissions_grid_sf, fao_regions) #sparse = FALSE)

emissions_zones_intersect <- st_intersects(emissions_grid_sf, fao_regions)

# add area column after intersection
emissions_zones$area_after <- st_area(emissions_zones)

# calculate area proportion
emissions_zones <- emissions_zones %>%
  mutate(area_prop = as.numeric(area_after/area_before)) %>%
  # multiply emissions by proportion
  mutate(co2_prop = total_co2 * area_prop) %>%
  select(-c(total_co2, area_prop, area_before, area_after))

# drop geometry
emissions_zones_no_geo <- emissions_zones %>%
  st_drop_geometry()

# -------- CHECKPOINT --------
#write_csv(emissions_zones_no_geo, file.path("/capstone/seamissions/checkpoint/emissions_zones_no_geo_vessel.csv"))
#st_write(emissions_zones, file.path("/capstone/seamissions/checkpoint/emissions_zones.gpkg"))
#write_csv(fao_catch_cleaned, file.path("/capstone/seamissions/checkpoint/fao_catch_cleaned.csv"))

emissions_zones_no_geo <- read_csv(file.path("/capstone/seamissions/checkpoint/emissions_zones_no_geo.csv"), show_col_types = FALSE)

#fao_catch <- read_csv(file.path("/capstone/seamissions/checkpoint/fao_catch_cleaned.csv"), show_col_types = FALSE)

```

```{r}
# take a look
test <- emissions_zones_no_geo %>%
  group_by(zone, year, flag) %>%
  summarise(co2 = sum(co2_prop))

# visualize co2 by region
emissions_zones %>%
  group_by(zone, year) %>%
  summarise(co2 = sum(co2_prop)) %>%
  group_by(zone) %>%
  summarise(mean_co2 = mean(co2)) %>%
  ggplot() +
    geom_sf(aes(fill = mean_co2)) +
  labs(fill = "Mean CO2 (mt)") +
  scale_fill_viridis_c(direction = -1) +
  theme_void()
```

## Decision points:

-   what attributes do we want to filter emissions for? do just `flag`, `year`, and `pollutants` make sense? maybe vessel class, but we cannot confidently say vessel class == species
-   what attributes do we want to filter fao catch data for? just `flag`, `year`, and `catch`?
-   how do we want to treat flag mis-match? merge flags?
-   how do we want to join catch and emissions? which is the primary dataset and how does that affect "lost" emissions

FAO flags **NOT** in emissions:

-   IOT (British Indian Ocean Territory)
-   CSK (Czechoslovakia)
-   ETH (Ethiopia)
-   GUM (Guam)
-   LSO (Lesotho)
-   MYT (Mayotte)
-   ANT (Netherlands Antilles)
-   NFK (Norfold Island)
-   BLM (Saint Barthelemy)
-   MAF (Saint Martin (french))
-   SRB (Serbia)
-   SCG (Serbia and Montenegro)
-   SXM (Sint Maarten (dutch))
-   SVK (Slovakia)
-   SSD (South Sudan)
-   SJM (Svalbard and Jan Mayen Islands)
-   TLS (Timor-Leste)
-   TKL (Tokelau)
-   UGA (Uganda)
-   SUN (Un. Sov. Soc. Rep.)
-   ESH (Western Sahara)
-   YUG (Yugoslavia SFR)
-   EAZ (United Republic of Tanzania, Zanzibar)

Emissions flags **NOT** in FAO:

-   UNK (unknown)
-   VAT (Holy See)
-   CXR (Christmas Island)
-   CCK (Cocos (Keeling) Islands)

### Proof of concept for Zone 31:

```{r}
# filter for zone 31
zone_31 <- emissions_zones %>%
  filter(zone == 31)

joined_31 <- left_join(zone_31, fao_catch, by = c("year", "flag", "zone")) %>%
  select(-c("lat_bin", "lon_bin"))

zone_31 <- emissions_zones_intersect %>%
  filter(zone == 31)

plot(st_geometry(zone_31))

plot(st)

# cut out grid cells that overlap with zone 31
#emissions_zone_31 <- st_intersection(emissions_grid_sf, zone_31) #sparse = FALSE)

# add area column
#emissions_zone_31$area_after <- st_area(emissions_zone_31)

# emissions_zone_31 <- zone_31 %>%
#   mutate(area_prop = as.numeric(area_after/area_before)) %>%
#   mutate(co2_prop = total_co2 * area_prop) %>%
#   select(-c(total_co2, area_prop, area_before, area_after))

# look at area breakdown to see if makes sense (mostly 1s)
# areas_31 <- emissions_zone_31 %>%
#   group_by(area_prop) %>%
#   summarise(count = n()) %>%
#   arrange(desc(count))

# Find which grid polygons intersect with the large polygon
overlapping_pixels <- st_intersects(emissions_grid_sf, zone_31, sparse = FALSE)

# Select grid polygons that have an intersection (TRUE values)
full_grid_31 <- emissions_grid_sf[overlapping_pixels, ]

# visualize the overlapping grid (st_intersects)
plot(st_geometry(zone_31))
plot(st_geometry(full_grid_31), add = TRUE)

# visualize the cut grid (st_intersection)
plot(st_geometry(zone_31))
plot(st_geometry(emissions_zone_31), add = TRUE)
```

```{r}
# -------- ZONE 31 FAO EMISSIONS --------
emissions_zone_31_no_geo <- emissions_zone_31 %>%
  st_drop_geometry()

fao_catch_cleaned_31 <- fao_catch_cleaned %>%
  filter(zone == 31)

# decision point
joined_31 <- left_join(emissions_zone_31_no_geo, fao_catch_cleaned_31, by = c("year", "flag", "zone"))

joined_31 <- left_join(fao_catch_cleaned_31, emissions_zone_31_no_geo, by = c("year", "flag", "zone"))
```

### Zone 41:

```{r}
# -------- 41 --------
# filter for zone 41
zone_31 <- fao_regions %>%
  filter(zone == 31)

emissions_zone_31 <- st_intersects(emissions_grid_sf, zone_31, sparse = FALSE) # sparse = FALSE)

plot(st_geometry(zone_31))
plot(st_geometry(emissions_zone_31_tion), add = TRUE)
plot(st_geometry(emissions_zone_31), add = TRUE)

emissions_zone_31 <- emissions_zone_31 %>%
  st_as_sf()

emissions_zone_31_tion <- st_intersection(emissions_grid_sf, zone_31)

sf2_intersections <- zone_31[unlist(emissions_zone_31), ]
plot(st_geometry(sf2_intersections))  # Now works



# cut out grid cells that overlap with zone 41
emissions_zone_41 <- st_intersects(emissions_grid_sf, zone_41) # sparse = FALSE)

plot(st_geometryemissions_zone_41)


# add area column
emissions_zone_41$area_after <- st_area(emissions_zone_41)

emissions_zone_41 <- emissions_zone_41 %>%
  mutate(area_prop = as.numeric(area_after/area_before)) %>%
  mutate(co2_prop = total_co2 * area_prop) %>%
  select(-c(total_co2, area_prop, area_before, area_after))

# look at area breakdown to see if makes sense (mostly 1s)
# areas_41 <- emissions_zone_41 %>%
#   group_by(area) %>%
#   summarise(count = n()) %>%
#   arrange(desc(count))

# Find which grid polygons intersect with the large polygon
overlapping_pixels <- st_intersects(emissions_grid_sf, zone_41, sparse = FALSE)

# Select grid polygons that have an intersection (TRUE values)
full_grid_41 <- emissions_grid_sf[overlapping_pixels, ]

# visualize the overlapping grid (st_intersects)
plot(st_geometry(zone_41))
plot(st_geometry(full_grid_41), add = TRUE)

# visualize the cut grid (st_intersection)
plot(st_geometry(zone_41))
plot(st_geometry(emissions_zone_41), add = TRUE)
```

```{r}
#| code-summary: Expand code

# Define emissions as sf object
# emissions_sf <- emissions %>%
#   st_as_sf(coords = c("lon_bin", "lat_bin"), crs = "54009") #4326 

# --- Prepare emissions and FAO data ----
emissions_sf <- emissions %>%
  # add lat and lon bin to preserve for merge 
  dplyr::mutate(lat = lat_bin, lon = lon_bin) %>%
  st_as_sf(coords = c("lon", "lat"), crs = "54009")  

# for sequoia
#fao_regions <- st_read(file.path("/home/emlab/projects/current-projects/ocean-ghg-fisheries/data/raw/world_fao_zones/world_fao_zones.shp"), crs = st_crs(emissions_sf)) #|> st_make_valid()

# for workbench
fao_regions <- st_read(file.path("/capstone/seamissions/data/fao_region_shapefile"), crs = st_crs(emissions_sf))
```

```{r}
# ---- Extract fao zone for emissions points ----

# Find which grid cell each point falls into
point_grid_intersections <- st_intersects(points, grid)

# Assign the grid cell ID to each point
points$grid_id <- sapply(point_grid_intersections, function(x) ifelse(length(x) > 0, x[1], NA))
```

1578367

```{r}

# subset sf for speed
subset_sf <- emissions_sf %>%
  filter(lat_bin %in% c(36:40)) %>%
  filter(lon_bin %in% c(-40:-45))

# ---- Step 1: Convert lat/long points to grid of polygons (1x1 degree cells) & make a key ----
emissions_grid <- subset_sf %>%
  # select lat & lon columns
  dplyr::select(lat_bin, lon_bin)

plot(st_geometry(emissions_grid))

#joined_grid <- st_join(emissions_grid, fao_regions, left = FALSE)
#unique(joined_grid$zone)

# check this step!
joined_grid <- st_join(emissions_grid, fao_regions, left = TRUE, join = st_intersection)
  
  # ---- Step 2: Find overlap between the grid cells and FAO zones using st_intersection ----
fao_intersections <- st_intersection(emissions_grid, fao_regions)
plot(fao_intersections)

# check it out
# Convert to sf to handle attributes and geometry together
fao_intersections <- st_as_sf(fao_intersections)

# Add area column to fao_intersections
joined_grid$area <- st_area(joined_gid)

# Now select the relevant columns (fao_zone, geometry, area)
fao_intersections <- fao_intersections %>%
  dplyr::select(fao_zone = zone, geometry, area)

# ---- Step 3: Join zones back to emissions ----
emissions_sf <- emissions_sf %>%
  st_join(fao_intersections) 
```

```{r}
emissions_sf <- emissions_sf %>%
  st_drop_geometry()
```

```{r}
# plot grid
plot(st_geometry(emissions_grid), main = "Emissions Grid")
```

## Side quest

```{r}
#| code-summary: Expand code
# Define emissions as sf object
emissions_sf_2 <- emissions %>%
  filter(lat_bin >= 0 & lat_bin <= 10) %>%
  filter(lon_bin >= 0 & lon_bin <= 10) %>%
  st_as_sf(coords = c("lon_bin", "lat_bin"), crs = "54009") # Change CRS later
  
# convert lat/long points to grid of polygons
emissions_grid_2 <- emissions_sf_2 %>%
  st_make_grid(cellsize = c(1,1),
               what = "polygons")

points_2 <- emissions %>%
  filter(lat_bin >= 0 & lat_bin <= 10) %>%
  filter(lon_bin >= 0 & lon_bin <= 10) %>%
  st_as_sf(coords = c("lon_bin", "lat_bin"), crs = "54009")


test <- emissions %>%
  filter(lat_bin == 8) %>%
  filter(lon_bin == 8)

# check it out
plot(st_geometry(emissions_grid_2), main = "Emissions Grid")
plot(st_geometry(points_2), add = TRUE)
```

#### Check CRS and projections

-   Check FAO and emissions grid, make sure they match and re-project if needed

```{r}

emissions_grid <- emissions_sf %>%
  st_make_grid(cellsize = c(1,1),
               what = "polygons")

st_crs(fao_regions) == st_crs(emissions_grid)

zone_31 <- fao_regions %>%
  filter(zone ==31)
#plot(zone_31)

#intersecting_points <- emissions_sf[st_intersects(emissions_sf, zone_31, sparse = FALSE), ]
zone_31_grid <- st_intersection(emissions_grid, zone_31, sparse = FALSE)

plot(zone_31_grid)

zone_31_grid$area <- st_area(zone_31_grid)

unique(zone_31_grid$area)
```

```{r}
# still mis-match?!
plot(zone_31, main = "FAO Zone 31", col = "gray90")
plot(zone_31_grid, add = TRUE)

st_crs(zone_31) == st_crs(zone_31_grid) #TRUE

```

# START HERE

-   it takes forever to run
-   resulting df saved to checkpoints to make imports easier and skip steps

### Assemble emissions data

```{r}
#| code-summary: Assemble emissions sf data
# read in emissions data from local checkpoint
#emissions <- read_csv(here("checkpoint", "emissions.csv"), show_col_types = FALSE)

# read in emissions data from server checkpoint
emissions <- read_csv(file.path("/capstone/seamissions/data/checkpoint/emissions.csv"))

# convert emissions to spatial object(?)
emissions_sf <- emissions %>%
  # add lat and lon bin to preserve for merge 
  dplyr::mutate(lat = lat_bin, lon = lon_bin) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 54009) # CRS not working for Carmen

# read in fao regions (assign same CRS as emissions_sf)
fao_regions <- st_read(file.path("/capstone/seamissions/data/fao_region_shapefile"), crs = st_crs(emissions_sf))

# create grid throughout extent of emissions_sf
emissions_grid <- emissions_sf %>%
  st_make_grid(cellsize = c(1,1),
               what = "polygons")

# convert the grid to an sf object
emissions_grid_sf <- st_sf(geometry = emissions_grid)

# spatially join emissions data to the grid
emissions_grid_sf <- st_join(emissions_grid_sf, emissions_sf, left = FALSE)

# If a grid cell has no emissions data, it will be removed when left = FALSE.
# If you want all grid cells, even empty ones, use left = TRUE.

# -------- CHECKPOINT --------
# save locally
st_write(emissions_grid_sf, here("checkpoint", "emissions_grid_sf.gpkg"))

# save to server
st_write(emissions_grid_sf, file.path("/capstone/seamissions/data/checkpoint/emissions_grid_sf.gpkg"))
```

#### Assemble FAO catch data

```{r}
#| code-summary: Assemble FAO catch data
# ---- FAO seafood production data ---
fao_capture_quantity <- read_csv(here::here("/capstone/seamissions/data/fao_seafood_production/Capture_Quantity.csv"), show_col_types = FALSE) |> clean_names()

fao_country_groups <- read_csv(here::here("/capstone/seamissions/data/fao_seafood_production/CL_FI_COUNTRY_GROUPS.csv"), show_col_types = FALSE)|> clean_names()

fao_species_groups <- read_csv(here::here("/capstone/seamissions/data/fao_seafood_production/CL_FI_SPECIES_GROUPS.csv"), show_col_types = FALSE)|> clean_names()

fao_symbol <- read_csv(here::here("/capstone/seamissions/data/fao_seafood_production/CL_FI_SYMBOL_SDMX.csv"), show_col_types = FALSE) |> clean_names()
fao_waterarea <- read_csv(here::here("/capstone/seamissions/data/fao_seafood_production/CL_FI_WATERAREA_GROUPS.csv"), show_col_types = FALSE)|> clean_names()

# --- Clean Data ----
fao_country_groups_clean <- fao_country_groups |>
  select(un_code, iso3_code, name_en, continent_group_en, official_name_en)

fao_species_groups_clean <- fao_species_groups |>
  select(x3a_code, taxonomic_code, identifier, scientific_name, contains("_en"))

fao_waterarea_clean <- fao_waterarea |>
  select(code, contains("_en"))

# ---- Left join capture and country groups ----
fao_capture_join <- left_join(fao_capture_quantity, fao_country_groups_clean, by = join_by(country_un_code == un_code)) 

fao_capture_join <- left_join(fao_capture_join, fao_species_groups_clean, by = join_by(iso3_code == x3a_code))

fao_capture_join <- left_join(fao_capture_join, fao_waterarea_clean, by = join_by(area_code == code))

# clean
fao_catch <- fao_capture_join %>%
  select(area_code,
         period,
         iso3_code,
         official_name_en,
         scientific_name,
         value) %>%
  rename(zone = area_code,
         year = period,
         flag = iso3_code,
         catch = value) %>%
  # filter for >2015
  filter(year >= 2015) %>%
  group_by(year, zone, flag, scientific_name, official_name_en) %>%
  summarise(total_catch = sum(catch, na.rm = TRUE)) %>%
  mutate(across(c(zone, year), as.factor))

# -------- CHECKPOINT --------
# save locally
write_csv(fao_catch, here("checkpoint", "fao_catch.csv"))

# save to server 
write_csv(fao_catch, file.path("/capstone/seamissions/data/checkpoint/fao_catch.csv"))
```

### This is proof of concept for zone 31:

```{r}
#| code-summary: Isolate Zone 31
# filter for zone 31
zone_31 <- fao_regions %>%
  filter(zone == 31)

# confirm crs match
st_crs(emissions_grid_sf) == st_crs(zone_31) # TRUE

# cut out grid cells that overlap with zone 31
emissions_zone_31 <- st_intersection(emissions_grid_sf, zone_31, sparse = FALSE)

# add area column
emissions_zone_31$area <- st_area(emissions_zone_31)

# look at area breakdown to see if makes sense (mostly 1s)
areas <- emissions_zone_31 %>%
  group_by(area) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# calculate co2 by pixel for zone 31 viz
zone_31_co2 <- emissions_zone_31 %>%
  group_by(lat_bin, lon_bin) %>%
  summarize(co2 = sum(total_co2))

# visualize co2 by pixel for zone 31
ggplot(zone_31_co2) +
  geom_sf(aes(fill = co2)) +
  theme_minimal()

# prep to merge with fao data
# prep_emissions_zone_31 <- emissions_zone_31 %>%
#   group_by(zone, year_month, flag) %>%
#   summarise(countries = n())

# -------- CHECKPOINT --------
# save locally
st_write(emissions_zone_31, here("checkpoint", "emissions_zone_31.gpkg"))

# save to server
st_write(emissions_zone_31, file.path("/capstone/seamissions/data/checkpoint/emissions_zone_31.gpkg"))
```

### This is proof of concept for zone 41:

```{r}
#| code-summary: Isolate Zone 31

#fao_regions <- st_cast(fao_regions, "POLYGON") %>%
 # st_make_valid()

mollweide_proj <- "+proj=moll +lon_0=0 +datum=WGS84 +units=m +no_defs"

emissions_grid_sf <- st_set_crs(emissions_grid_sf, mollweide_proj)

emissions_grid_sf <- st_set_crs(emissions_grid_sf, 8857)

# remove warning?
emissions_grid_sf <- st_make_valid(emissions_grid_sf)

fao_regions <- st_read(file.path("/capstone/seamissions/data/fao_region_shapefile"), crs = st_crs(emissions_grid_sf))

# filter for zone 31
zone_41 <- fao_regions %>%
  filter(zone == 41)

# confirm crs match
st_crs(emissions_grid_sf) == st_crs(zone_41) # TRUE

# cut out grid cells that overlap with zone 31
emissions_zone_41 <- st_intersection(emissions_grid_sf, zone_41, sparse = FALSE)

# add area column
emissions_zone_41$area <- st_area(emissions_zone_41)

# look at area breakdown to see if makes sense (mostly 1s)
areas <- emissions_zone_41 %>%
  group_by(area) %>%
  summarise(count = n()) %>%
  arrange(desc(count))


st_write(emissions_grid_sf, here("checkpoint", "emissions_grid_sf.gpkg"))

st_write(emissions_grid_sf, file.path("/capstone/seamissions/data/checkpoint/emissions_grid_sf.gpkg"))

plot(emissions_zone_41)


# visualize co2 by pixel for zone 41
# ggplot(emissions_zone_41) +
#   geom_sf() +
#   theme_minimal()

# -------- CHECKPOINT --------
# save locally
st_write(emissions_zone_41, here("checkpoint", "emissions_zone_41.gpkg"))

# save to server
st_write(emissions_zone_41, file.path("/capstone/seamissions/data/checkpoint/emissions_zone_41.gpkg"))
```

### Full join (takes SO much time)

```{r}
#| code-summary: Join all data
# -------- FULL join --------

# clean emissions_grid_sf first??

# join
emissions_regions <- st_intersection(emissions_grid_sf, fao_regions, sparse = FALSE)

# add area column
emissions_regions$area <- st_area(emissions_regions)

# -------- CHECKPOINT --------
#save locally
st_write(emissions_regions, here("checkpoint", "emissions_regions.gpkg"))

# save to server
st_write(emissions_regions, file.path("/capstone/seamissions/data/checkpoint/emissions_regions.gpkg"))
```

### Attempt to merge emissions and FAO catch data

*make sure flags match in FAO and GFW datasets*

FAO flags **NOT** in emissions:

-   IOT (British Indian Ocean Territory)
-   CSK (Czechoslovakia)
-   ETH (Ethiopia)
-   GUM (Guam)
-   LSO (Lesotho)
-   MYT (Mayotte)
-   ANT (Netherlands Antilles)
-   NFK (Norfold Island)
-   BLM (Saint Barthelemy)
-   MAF (Saint Martin (french))
-   SRB (Serbia)
-   SCG (Serbia and Montenegro)
-   SXM (Sint Maarten (dutch))
-   SVK (Slovakia)
-   SSD (South Sudan)
-   SJM (Svalbard and Jan Mayen Islands)
-   TLS (Timor-Leste)
-   TKL (Tokelau)
-   UGA (Uganda)
-   SUN (Un. Sov. Soc. Rep.)
-   ESH (Western Sahara)
-   YUG (Yugoslavia SFR)
-   EAZ (United Republic of Tanzania, Zanzibar)

Emissions flags **NOT** in FAO:

-   UNK (unknown)
-   VAT (Holy See)
-   CXR (Christmas Island)
-   CCK (Cocos (Keeling) Islands)

```{r}
#| code-summary: Another approach

# filter/group emissions sf data to reduce compute time
emissions_sf_2 <- emissions_sf %>%
  #filter(lat_bin > 0) %>% # remove later
  #filter(lon_bin > 0) %>% # remove later
  mutate(year = as.integer(substr(year_month, 1, 4))) %>%
  select(-year_month) %>%
  group_by(year, lat_bin, lon_bin, flag) %>%
  summarise(total_co2 = sum(total_co2, na.rm = TRUE))

# check for lost emissions
sum(emissions_grid_sf$total_co2, na.rm = TRUE) == sum(emissions_grid_sf_merge$total_co2, na.rm = TRUE)

# assign FAO regions to filtered/grouped df
emissions_regions <- st_intersection(emissions_grid_sf_2, fao_regions, sparse = FALSE)

# add area column
emissions_regions$area <- st_area(emissions_regions)

# spatially join emissions data to the grid
emissions_grid_sf_2 <- st_join(emissions_regions, emissions_sf_2, left = FALSE)

emissions_regions <- emissions_regions %>%
  mutate(prop_total_co2 = (area * total_co2)) %>%
  select(-c("area", "total_co2", "lat_bin", "lon_bin")) %>%
  st_drop_geometry()

# merge emissions data with FAO catch data
catch_emissions <- left_join(emissions_regions, fao_catch, by = c("flag", "year"))

# -------- CHECKPOINT --------
# save locally
write_csv(catch_emissions, here("checkpoint", "catch_emissions.csv"))

# save to server
write_csv(catch_emissions, file.path("/capstone/seamissions/data/checkpoint/catch_emissions.csv"))
```

### Attempt to merge emissions with FAO catch data for zone 31

```{r}
#| code-summary: Merge emissions and FAO catch data
# flags in fao catch but not emissions for zone 31
diff_flag <- anti_join(fao_catch_31, emissions_zone_31, by = "flag") 

# replace year_month with year
emissions_zone_31 <- emissions_zone_31 %>%
  mutate(year = as.integer(substr(year_month, 1, 4))) %>%
  select(-year_month) %>%
  mutate(prop_total_co2 = (area * total_co2)) %>%
  select(-c("area", "total_co2", "lat_bin", "lon_bin")) %>%
  st_drop_geometry()

# join by zone, year, and flag
catch_emissions_zone_31 <- left_join(emissions_zone_31, fao_catch, by = c("zone", "year", "flag")) #relationship = "many-to-many")

# drop geometry
#catch_emissions_zone_31_no_geo <- catch_emissions_zone_31 %>%
  #st_drop_geometry()

# ------- CHECKPOINT --------
st_write(catch_emissions_zone_31, here("checkpoint", "catch_emissions_zone_31.gpkg"))

write_csv(catch_emissions_zone_31_no_geo, here("checkpoint", "catch_emissions_zone_31_no_geo.csv"))
```

Flags in FAO catch but **not** emissions (for Zone 31):

-   AIA (Anguilla)
-   BHS (Bahamas)
-   BRB (Barbados)
-   CUB (Cuba)
-   DMA (Dominica)
-   GUF (French Guiana)
-   GLP (Guadeloupe)
-   GTM (Guatemala)
-   GUY (Guyana)
-   HND (Honduras)
-   ITA (Italy)
-   MTQ (Martinique)
-   MSR (Montserrat)
-   ANT (Netherlands Antilles)
-   PHL (Philippines)
-   PRI (Puerto Rico)
-   BLM (Saint Barthelemy)
-   LCA (Saint Lucia)
-   MAF (Saint Martin (french))
-   SXM (Sint Maarten (dutch))
-   SUR (Suriname)
-   SUN (Un. Sov. Soc. Rep.)
-   VIR (US Virgin Islands)

#### Explore

```{r}
unique(fao_zone_31_merged_no_geo$scientific_name)


sample <- fao_zone_31_merged_no_geo %>%
  filter(scientific_name == "Panulirus spp") # spiny lobster

sample_viz <- sample %>%
  group_by(year, flag) %>%
  summarise(sum_co2 = sum(total_co2),
            sum_counts = sum(count)) %>%
  mutate(cost = sum_co2/sum_counts)

sample_2 <- fao_zone_31_merged_no_geo %>%
  filter(scientific_name == "Setipinna phasa") # 

sample_2_viz <- sample_2 %>%
  group_by(year, flag) %>%
  summarise(sum_co2 = sum(total_co2),
            sum_counts = sum(count)) %>%
  mutate(cost = sum_co2/sum_counts)

top <- fao_zone_31_merged_no_geo %>%
  group_by(flag, scientific_name) %>%
  summarise(counts = n()) %>%
  #filter(scientific_name == "Austrolabrus maculatus") %>%
  arrange(desc(counts))

sample_3 <- fao_zone_31_merged_no_geo %>%
  filter(scientific_name == "Austrolabrus maculatus") # black spotted parrotfish?

sample_3_viz <- sample_3 %>%
  group_by(year, flag) %>%
  summarise(sum_co2 = sum(total_co2),
            sum_counts = sum(count)) %>%
  mutate(cost = sum_co2/sum_counts)  


unique(fao_zone_31_merged_no_geo$flag)
```

-   Assign FAO regions to emissions
-   Assign majority pixel to region as a test then, assign by proportion

```{r}
# View the filtered results
#print(multiple_regions)

region_31 <- fao_regions %>%
  filter(zone == 31)

region_34 <- fao_regions %>%
  filter(zone == 34)

region_77 <- fao_regions %>%
  filter(zone == 77)

region_21 <- fao_regions %>%
  filter(zone == 21)

region_27 <- fao_regions %>%
  filter(zone == 27)


grid_31 <- multiple_regions %>%
  filter(str_detect(region, "31"))

grid_31_27 <- grid_31 %>%
  filter(str_detect(region, "27"))

#[1] "34,31"    "77,31"    "31,21"    "31,21,27" "34,31,27"

plot(st_geometry(region_31), col = "cornflowerblue")
plot(st_geometry(region_34), col = "lightblue", add = TRUE)
plot(st_geometry(region_77), col = "navyblue", add = TRUE)
plot(st_geometry(region_21), col = "lightgreen", add = TRUE)
plot(st_geometry(region_27), col = "darkgreen", add = TRUE)
#plot(st_geometry(grid_31), add = TRUE)
plot(st_geometry(grid_31_27))


bb <- st_bbox(grid_31_27)

plot(st_geometry(region_31), col = "cornflowerblue", xlim = bb[c(1,3)], ylim = bb[c(2,4)])
plot(st_geometry(region_27), col = "lightgreen", xlim = bb[c(1,3)], ylim = bb[c(2,4)], add = TRUE)
plot(st_geometry(region_34), col = "lightblue", xlim = bb[c(1,3)], ylim = bb[c(2,4)], add = TRUE)
plot(st_geometry(grid_31_27), add = TRUE)


grid_77 <- grid_31 %>%
  filter(str_detect(region, "77"))

bb <- st_bbox(grid_77)

plot(st_geometry(region_77), col = "lightgreen", xlim = bb[c(1,3)], ylim = bb[c(2,4)])
plot(st_geometry(region_31), col = "cornflowerblue", xlim = bb[c(1,3)], ylim = bb[c(2,4)], add = TRUE)
plot(st_geometry(grid_77), add = TRUE)



region_81 <- fao_regions %>%
  filter(zone == 81)

region_57 <- fao_regions %>%
  filter(zone == 57)


grid_81 <- multiple_regions %>%
  filter(str_detect(region, "81"))

grid_81 <- emissions_zones %>%
  filter(region == 81)

region_71 <- fao_regions %>%
  filter(zone ==71)



# Bounding box in Equal Earth projection (EPSG:8857)
australia_nz_bbox <- c(
  xmin = 900000,   # Covers Western Australia
  xmax = 2300000,  # Covers New Zealand's easternmost islands
  ymin = -6500000, # Covers Tasmania and southernmost NZ
  ymax = -2500000  # Covers northern Australia
)

bb <- st_bbox(region_57)


new_xmin <- 11000000  # Move western boundary even closer to Australia's east coast
new_xmax <- 12300000  # Move eastern boundary even closer to New Zealand
bbox_adjusted <- c(xmin = new_xmin, ymin = -6005522, xmax = new_xmax, ymax = 2928172)

plot(st_geometry(region_81), col = "cornflowerblue", xlim = bbox_adjusted[c(1,3)], ylim = bbox_adjusted[c(2,4)])
plot(st_geometry(region_57), col = "lightgreen", xlim = bbox_adjusted[c(1,3)], ylim = bbox_adjusted[c(2,4)], add = TRUE)
plot(st_geometry(region_71), col = "lavender", xlim = bbox_adjusted[c(1,3)], ylim = bbox_adjusted[c(2,4)], add = TRUE)
#plot(st_geometry(grid_81), add = TRUE)


```

```{r}
# cut out grid cells that overlap with zones (st_intersection)
emissions_zones <- st_intersection(emissions_grid_sf, fao_regions) #sparse = FALSE)

# add area column after intersection
emissions_zones$area_after <- st_area(emissions_zones)

# calculate area proportion
emissions_zones <- emissions_zones %>%
  mutate(area_prop = as.numeric(area_after/area_before)) %>%
  # multiply emissions by proportion
  mutate(co2_prop = total_co2 * area_prop) %>%
  select(-c(total_co2, area_prop, area_before, area_after))

# drop geometry
emissions_zones_no_geo <- emissions_zones %>%
  st_drop_geometry()

```
