---
title: "Option 1: FAO"
author: "Carmen Hoyt"
date: last-modified
format:
  html:
      toc: true
      code-fold: true
editor: visual
execute:
  warning: false
  messages: false
editor_options: 
  chunk_output_type: console
---

# Methods:

- For each year, calculate total **non-broadcasting** emissions for each FAO region: 

`fao_summary_non_broadcasting`

- Then divy that up to *all* fisheries that report catch in that region (proportionally by the amount of catch in each fishery). 

TABLE 1: `total_region_non_broad_allocation`

- For each year, calculate total **broadcasting** emissions, by country, for each FAO region. 

TABLE 2: `emissions_partitioned`

- Full join the allocated non-broadcasting emissions with the broadcasting emissions. 

TABLE 3: `total_region_emissions`

Assumptions:

- Within an FAO region, the non-broadcasting emissions are spread proportionally across *all* fisheries and countries within that region *by catch*.

## Load Packages

```{r}
#| code-summary: Load Packages
library(tidyverse)
library(janitor)
library(here)
#library(dplyr)
```

## Load Data

```{r}
#| code-summary: Load Data
emissions_partitioned_grouped <- read_csv(file.path("/capstone/seamissions/checkpoint/emissions_partitioned_grouped.csv"), show_col_types = FALSE)

fao_catch <- read_csv(file.path("/capstone/seamissions/checkpoint/fao_catch.csv"), show_col_types = FALSE)
```

## Partition Emissions by Region and Combine

```{r}
#| code-summary: Distribute Emissions
# Turn off scientific notation
options(scipen=999)

# Define regions
regions <- c(18, 21, 27, 31, 34, 47, 48, 51, 57, 58, 61, 67, 71, 77, 81, 87, 88)

# Create empty vector to append Region #
option_1 <- vector("list", length = length(regions))

# Partition emissions
for (i in seq_along(regions)){
  
  # Make list of flags in FAO catch data by Region #
  # fao_catch_flags <- fao_catch %>%
  #   filter(zone == regions[i])
  # 
  # fao_catch_flags <- unique(fao_catch_flags$flag)
  # 
  # # Make a list of flags in emissions data by Region #
  # emissions_flags <- emissions_partitioned_grouped %>%
  #   filter(zone == regions[i])
  # 
  # emissions_flags <- unique(emissions_flags$flag) # TABLE 2: flags with BROADCASTING emissions
  # 
  # # Make a list of flags with non-broadcasting emissions ONLY by Region # (those flags in FAO catch but not in emissions)
  # non_broad_only_flags <- setdiff(fao_catch_flags, emissions_flags)
  # 
  # Make a list of broadcasting only flags? check in final df?
  
 # -------- TABLE 1 --------
  # 1.1 All flags reporting FAO catch in Region #
  total_region_catch <- fao_catch %>%
    filter(zone == regions[i]) %>%
    group_by(zone, year, flag, species) %>%
    summarise(country_total_tons_by_species = sum(fao_catch_tons, na.rm = TRUE), .groups = "drop") %>%
    ungroup() %>%
    group_by(zone, year) %>%
    mutate(region_total_tons = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
    mutate(prop_fao_catch = country_total_tons_by_species/region_total_tons) %>%
    select(-region_total_tons) %>%
    filter(country_total_tons_by_species > 0)
  
  # 1.2 Summarise non-broadcasting emissions by Region #
  fao_summary_non_broadcasting <- emissions_partitioned_grouped %>%
    # Filter for Region # (to account for emissions later)
    filter(zone == regions[i]) %>%
    # Select non-broadcasting emissions
    filter(flag == "DARK") %>%
    # Rename columns for non-broadcasting emissions
    mutate(non_broad_co2_mt = co2,
           non_broad_ch4_mt = ch4,
           non_broad_n2o_mt = n2o,
           non_broad_nox_mt = nox,
           non_broad_sox_mt = sox,
           non_broad_co_mt = co,
           non_broad_vocs_mt = vocs,
           non_broad_pm2_5_mt = pm2_5,
           non_broad_pm10_mt = pm10) %>%
    # Keep only desired columns
    select(zone, 
           year, 
           non_broad_co2_mt,
           non_broad_ch4_mt,
           non_broad_n2o_mt,
           non_broad_nox_mt,
           non_broad_sox_mt,
           non_broad_co_mt,
           non_broad_vocs_mt,
           non_broad_pm2_5_mt,
           non_broad_pm10_mt) %>%
    # FAO data doesn't go past 2022
    filter(year <= 2022) 
  
  # TABLE 1: Allocate non-broadcasting emissions to FAO reporting countries by Region #
  total_region_non_broad_allocation <- left_join(total_region_catch, fao_summary_non_broadcasting, by = c("zone", "year")) %>%
    mutate(dist_non_broad_co2_mt = prop_fao_catch * non_broad_co2_mt,
           dist_non_broad_ch4_mt = prop_fao_catch * non_broad_ch4_mt,
           dist_non_broad_n2o_mt = prop_fao_catch * non_broad_n2o_mt,
           dist_non_broad_nox_mt = prop_fao_catch * non_broad_nox_mt,
           dist_non_broad_sox_mt = prop_fao_catch * non_broad_sox_mt,
           dist_non_broad_co_mt = prop_fao_catch * non_broad_co_mt,
           dist_non_broad_vocs_mt = prop_fao_catch * non_broad_vocs_mt,
           dist_non_broad_pm2_5_mt = prop_fao_catch * non_broad_pm2_5_mt,
           dist_non_broad_pm10_mt = prop_fao_catch * non_broad_pm10_mt) %>%
    select(-country_total_tons_by_species)
  
  # Account for emissions transfer in TABLE 1
  before <- sum(fao_summary_non_broadcasting$non_broad_co2_mt, na.rm = TRUE) 
  after <- sum(total_region_non_broad_allocation$dist_non_broad_co2_mt, na.rm = TRUE) 
  
  percent_diff <- ((after-before)/before)*100
  
  # Warning if emissions loss is greater than 0.001%
  if(percent_diff > 0.001) {
    stop(paste0("Greater than 0.001% of non-broadcasting emissions lost during allocation to FAO reporting countries in Region ", regions[i], ". Check Table 1."))
  }
  
  # -------- TABLE 2 --------
  # 2.1: All flags with broadcasting emissions by Region #
  emissions_partitioned <- emissions_partitioned_grouped %>%
    # Filter for Region #
    filter(zone == regions[i]) %>%
    # Remove non-broadcasting emissions
    filter(!flag == "DARK") %>%
    mutate(broad_co2_mt = co2,
           broad_ch4_mt = ch4,
           broad_n2o_mt = n2o,
           broad_nox_mt = nox,
           broad_sox_mt = sox,
           broad_co_mt = co,
           broad_vocs_mt = vocs,
           broad_pm2_5_mt = pm2_5,
           broad_pm10_mt = pm10) %>%
    select(-c(co2, ch4, n2o, nox, sox, co, vocs, pm2_5, pm10))
  
  # 2.2 Distribute broadcasting for countries with catch ASSUMPTION: 1 ton = same emissions for all species
  total_flag_catch <- fao_catch %>%
    filter(zone == regions[i]) %>%
    group_by(zone, year, flag, species) %>%
    summarise(country_total_tons_by_species = sum(fao_catch_tons, na.rm = TRUE), .groups = "drop") %>%
    ungroup() %>%
    group_by(zone, year, flag) %>%
    mutate(country_total_tons = sum(country_total_tons_by_species, na.rm = TRUE)) %>%
    mutate(prop_species_catch = country_total_tons_by_species/country_total_tons) %>%
    filter(country_total_tons_by_species > 0) %>%
    select(-c(country_total_tons)) %>% #country_total_tons_by_species
    filter(!is.na(prop_species_catch))
  
  # TABLE 2
  total_region_broad_allocation <- full_join(total_flag_catch, emissions_partitioned, by = c("zone", "year", "flag")) %>%
    mutate(dist_broad_co2_mt = ifelse(is.na(prop_species_catch), broad_co2_mt, prop_species_catch * broad_co2_mt),
           dist_broad_ch4_mt = ifelse(is.na(prop_species_catch), broad_ch4_mt, prop_species_catch * broad_ch4_mt),
           dist_broad_n2o_mt = ifelse(is.na(prop_species_catch), broad_n2o_mt, prop_species_catch * broad_n2o_mt),
           dist_broad_nox_mt = ifelse(is.na(prop_species_catch), broad_nox_mt, prop_species_catch * broad_nox_mt),
           dist_broad_sox_mt = ifelse(is.na(prop_species_catch), broad_sox_mt, prop_species_catch * broad_sox_mt),
           dist_broad_co_mt = ifelse(is.na(prop_species_catch), broad_co_mt, prop_species_catch * broad_co_mt),
           dist_broad_vocs_mt = ifelse(is.na(prop_species_catch), broad_vocs_mt, prop_species_catch * broad_vocs_mt),
           dist_broad_pm2_5_mt = ifelse(is.na(prop_species_catch), broad_pm2_5_mt, prop_species_catch * broad_pm2_5_mt),
           dist_broad_pm10_mt = ifelse(is.na(prop_species_catch), broad_pm10_mt, prop_species_catch * broad_pm10_mt))
  
  # -------- TABLE 3 --------
  # TABLE 3: Full join Tables 1 and 2 (to not lose flags with non-broad emissions but no broad emissions)
  total_region_emissions <- full_join(total_region_broad_allocation, total_region_non_broad_allocation, by = c("zone", "year", "flag", "species")) %>%
    #filter(zone == 18) %>%
    # Sum broadcasting and non-broadcasting emissions
    mutate(total_co2_mt = rowSums(across(c(dist_broad_co2_mt, dist_non_broad_co2_mt)), na.rm = TRUE),
           total_ch4_mt = rowSums(across(c(dist_broad_ch4_mt, dist_non_broad_ch4_mt)), na.rm = TRUE),
           total_n2o_mt = rowSums(across(c(dist_broad_n2o_mt, dist_non_broad_n2o_mt)), na.rm = TRUE),
           total_nox_mt = rowSums(across(c(dist_broad_nox_mt, dist_non_broad_nox_mt)), na.rm = TRUE),
           total_sox_mt = rowSums(across(c(dist_broad_sox_mt, dist_non_broad_sox_mt)), na.rm = TRUE),
           total_co_mt = rowSums(across(c(dist_broad_co_mt, dist_non_broad_co_mt)), na.rm = TRUE),
           total_vocs_mt = rowSums(across(c(dist_broad_vocs_mt, dist_non_broad_vocs_mt)), na.rm = TRUE),
           total_pm2_5_mt = rowSums(across(c(dist_broad_pm2_5_mt, dist_non_broad_pm2_5_mt)), na.rm = TRUE),
           total_pm10_mt = rowSums(across(c(dist_broad_pm10_mt, dist_non_broad_pm10_mt)), na.rm = TRUE))
  
  # Keep desired columns
  total_region_emissions <- total_region_emissions %>%
    select(year,
           zone,
           flag,
           species,
           c(colnames(total_region_emissions)[str_detect(colnames(total_region_emissions), "total")])
    )
  
  # BUILD SPECIES TABLE for zone, year, flag, species, catch_tons
  
  # Account for emissions after full join
  before <- sum(emissions_partitioned$broad_co2_mt, na.rm = TRUE) + sum(fao_summary_non_broadcasting$non_broad_co2_mt, na.rm = TRUE)
  after <- sum(total_region_emissions$total_co2_mt, na.rm = TRUE) 
  
  percent_diff <- ((after-before)/before)*100
  
  # Warning if emissions loss is greater than 0.001%
  if(percent_diff > 0.001) {
    stop(paste0("Greater than 0.001% of emissions lost during full join of broadcasting and non-broadcasting in Region ", regions[i], ". Check Table 3."))
  }
  
  # Append to vector
  option_1[[i]] <- total_region_emissions
}

# Full emissions dataset
full_emissions_fao <- bind_rows(option_1)
```

### Checkpoint

```{r}
#| code-summary: Checkpoint

# Save FAO emissions dataset to workbench-2
write_csv(full_emissions_fao, file.path("/capstone/seamissions/checkpoint/full_emissions_fao.csv"))

# Read in FAO emissions dataset
#full_emissions_fao <- read_csv(file.path("/capstone/seamissions/checkpoint/full_emissions_fao.csv"), show_col_types = FALSE)
```
